### 2021-12-13 paper/paper.tex

 \documentclass[11pt,onecolumn]{article}
 \usepackage[margin=0.5in]{geometry}

\usepackage[ruled,vlined]{algorithm2e}
\usepackage{amsmath}
\usepackage{xspace}
\usepackage[binary-units=true]{siunitx}
\usepackage{ulem}
%\usepackage{censor}
\usepackage[table]{xcolor}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{tabularx}

\usepackage{subcaption}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{listings}
\usepackage{dingbat}
\usepackage{mathtools}

\hypersetup{
    colorlinks=true,
    linkcolor=black,
    citecolor=blue,
    filecolor=black,
    urlcolor=blue}

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\makeatletter
  \def\footnoterule{\kern-3\p@
  \hrule \@width 2in \kern 2.6\p@} % the \hrule is .4pt high
\makeatother

\begin{document}

\newcommand{\fslspm}{FSL-SPM\xspace}
\newcommand{\fslafni}{FSL-AFNI\xspace}
\newcommand{\afnispm}{AFNI-SPM\xspace}
\newcommand{\tristan}[1]{\color{orange}\textbf{From Tristan:} #1\color{black}\xspace}
\newcommand{\camille}[1]{\color{blue}\textbf{From Camille:} #1\color{black}\xspace}
\newcommand{\ali}[2]{\color{green}\textbf{Ali:} #1\color{black}\xspace}
\newcommand{\discuss}[1]{\uwave{#1}}
\newcommand{\gk}[1]{\color{purple}#1 \textbf{-GK}\color{black}\xspace}
\newcommand{\yohan}[1]{\color{cyan!75!black} \textbf{Yohan:} #1\color{black}\xspace}
\newcommand{\yohanmod}[1]{\color{cyan!75!black} \sout{#1}\color{black}\xspace}

\title{Comparing variability across fMRI analysis software packages and numerical errors}

\author{Ali Salari$^1$, Yohan Chatelain$^1$, Alexander Bowring$^2$, Camille Maumet$^3$, Gregory Kiar$^4$, Tristan Glatard$^1$\\
  \vspace*{0.1cm}\\
  $^1$ Department of Computer-Science and Software Engineering, Concordia University, Montreal, Canada\\
  $^2$ Li Ka Shing Centre for Health Information and Discovery, Nuffield Department of Population Health,\\ Big Data Institute, University of Oxford, Oxford, UK\\
  $^3$ Inria, Univ Rennes, CNRS, Inserm, IRISA UMR 6074, Empenn ERL U 1228, Rennes, France\\
  $^4$ Center for the Developing Brain, Child Mind Institute, New York, NY, USA}

\maketitle
\begin{abstract}
  Variability has been broadly observed in functional MRI analyses across both
  differences in analytic tools and numerical instabilities. However,
  the relationship between numerical (within-tool) and between-tool variabilities for
  given analysis is unclear. We extended a previous comparison
  of fMRI analysis software libraries (namely FSL, AFNI, and SPM) and relate the observed differences to the numerical
  variability observed using Monte-Carlo arithmetic. We found that
  between-tool variability was consistently larger than numerical variability in group-level analyses,
  whereas numerical variability approached between-tool variability in some subject-level analyses.
  Interestingly, between-tool variability and
  machine error appeared moderately correlated ($r > 0.5$) in all tested conditions.
  Finally, we found that numerical variability had effects of similar magnitude to between-tool
  variability when errors were introduced at the precision of 17 bits --- a
  precision level found in commonly-used libraries \gk{What does this mean? They only have precision of 17, or they use data that is 17 bits?}. \yohan{agree with Greg, it is unclear and a strong assertion!} \camille{It is not entirely clear to me how this relates to the previous conclusions that BT variability was consistently larger than machine error} Our findings motivate the
  continued investigation of numerical instability in neuroimaging, and position it
  as a possible proxy for studying arbitrary between-tool variations.
\end{abstract}

\section{Introduction}

Software variability has been highlighted as an important source of error
in fMRI analysis, originating from mainly two factors: between-tool
variability denotes differences in the outcomes of different
software toolboxes --- such as FSL, AFNI and SPM --- when analyzing the same dataset
using similar methods \gk{this sentence says "two factors" but then only mentioned between tool. Also, it is too long}. \yohan{What do you mean by methods? Numerical schemes? If so, are we sure that theses tools are using the same numerical schemes?} \camille{I also had highlighed the 2 here, I think we could use "multiple factors" --> in practice we could also add different types of within-software variability (in the form of changing algorithms and/or parameters for a given step)} Numerical variability, in contrast, denotes
differences in outcomes of the same analytical toolbox
executed in different environments – such as different operating systems or
different hardware platforms \gk{you mention these two examples of numerical variability, but you use neither in this paper. Why not mention the one you use?}. This manuscript investigates the extent to
which both types of variability relate to each other.

Numerical variability stems from errors introduced in floating point
computations by \yohanmod{numerical} rounding, cancellation, and absorption~\cite{muller2018handbook} \gk{only? What about poor conditioning + random effects + noise in data?}. In
practice, the occurrence of such errors is affected by variations in
hardware, software dependencies, compilation, parallelization, and most
likely other factors \gk{what does "most likely other factors" mean?}. In the best case, the impact of such errors and the induced
variations is in the range of \yohanmod{1} \yohan{1/2 ulp for round-to-nearest} unit in the last place (ulp) --- also known
as machine error, the difference between two successive floating-point
numbers. However, when computations depend on imprecise libraries,
numerical error may vary by several ulps. For instance, the \yohan{maximal} numerical error
of the complementary error function (erfc) is larger than 3 ulps in several
mathematical libraries~\cite{zimmermann:hal-03141101}, implying that an update of this function could introduce
numerical \yohanmod{perturbations} \yohan{differences} larger than 3 ulps. Due to numerical instability,
machine error may amplify during its propagation through analyses \yohan{do you mean calculations?},
resulting in considerable loss of precision in final results. In fMRI analysis,
numerical variability produced by Linux updates has been shown to result in Dice
coefficients ranging from 0.0 to 1.0 for ICA components produced by FSL
MELODIC~\cite{Glatard2015} \gk{there isn't really context for what this result means, yet. Is this a lot? Dice hasn't been explained. Also, how are ICA components (weighted vectors) compared via Dice (a measure for binary image/vector comparison)? }.

Between-tool variability is not a surprising concept: different programs
executed on the same data may obviously produce different results \gk{even when doing conceptually the same thing? Why? Why is this obvious?}. However,
in fMRI analysis, programs all aim to estimate brain activity from measured
variations in the BOLD signal, which is not expected to be measurably
impacted by computational artifacts \gk{Why wouldn't it be? Especially if you state that it isn't surprising for tools to lead to different results, how come you expect fMRI to not suffer from this?}. \yohan{+1} Therefore, the observation reported
in~\cite{bowring2019exploring} that the three main fMRI analysis toolboxes
may result in marked analytical differences (Dice coefficients ranging from
0.000 to 0.684 \camille{Apologies I don't have the numbers on the top of my head but can I confirm taht those were taken from the erratum that was published after the paper?} for thresholded activation maps), has raised substantial
concern. In a follow-up study~\cite{bowring2021isolating}, the same authors
isolated the analytical components where the pipelines diverge,
highlighting a major impact of the signal model and suggesting that the
root-cause for between-tool variability depends on the analysis design and
task paradigm \gk{If you state this finding, you need to unpack what it means a little bit. What part of analysis design? By task paradigm do you essentially mean the expected effect-size for the task that individuals are performing within the scanner?} \camille{would be worth adding here that the identified main sources of variations were different across the 3 studies}. Further, the work in~\cite{Li2021.12.01.470790} found
moderate agreement between results produced by five independent fMRI
analysis pipelines and analyzed in details the components that led to these
variations \gk{This sentence feels inconsistent and counter to the previous, and also doesn't tell us their findings; what components led to variations?}. Between-tool variability is supposedly an important determinant
of the substantial analytical variability recently observed in
fMRI~\cite{botvinik2020variability} \gk{arguably, one of the larger issues that the NARPS result showed us was that people construct and test ``equivalent'' hypotheses differently, so despite highly similar data they will arrive at different conclusions}.

We investigate the relationship between numerical and between-tool
variability through two main questions: (1) what is the relative magnitude of
within-tool (numerical) variability and between-tool variability, and (2) is there an association between within- and between-tool variability. The first question determines the extent to which each variability
source may impact current fMRI results, and which one should be addressed or leveraged
in priority. The second question may provide deeper insights on the origins
of software variability. Conceptually, both types of variability may
reflect a single solution space shaped by local minima and
ill-conditioning \gk{this feels a bit clumsily worded, but I don't immediately have a rewrite that comes to mind }, which could be explored through numerical or software
perturbations.

\section{Materials and Methods}

\subsection{fMRI analysis \& Dataset}

We replicated the analysis described as study `ds000001'
in~\cite{schonberg2012decreasing}, relying on the data publicly available
in OpenNeuro at \url{https://openneuro.org/datasets/ds000001} and using
three widely-used software packages for fMRI data processing, namely FMRIB
Software Library (FSL)~\cite{jenkinson2012fsl}, Analysis of Functional
NeuroImages (AFNI)~\cite{cox1996afni}, and Statistical Parametric
Mapping (SPM)~\cite{penny2011statistical}. We selected this dataset because
comparable analysis pipelines implemented in FSL, AFNI and SPM were already
publicly available and extensively described in~\cite{bowring2019exploring}.
Furthermore, the work in~\cite{bowring2019exploring} already evaluated the
effect of tool variability for this dataset, which we intended to
extend with the present quantification of numerical variability.

In the selected study, 16 healthy adult subjects participated in the
balloon analog risk task~\cite{lejuez2002evaluation} to measure risk-taking
behavior over three scanning sessions~\cite{schonberg2012decreasing}. We
reused the preprocessing, first-level, and second-level analyses
implemented by~\cite{bowring2019exploring} consistently across all three
software packages. Table~\ref{table:pipeline-steps}, adapted from~\cite{bowring2019exploring},
summarizes the analytical steps in each
pipeline.

%%%%%%%%%% Summary of statstics %%%%%%%%
\setlength{\tabcolsep}{4pt}
\begin{table}[h]
  \centering
  \begin{tabular}{|c|l|c|c|c|}
    \hline
    %        \multirow{2}{*}{} & \multicolumn{1}{c}{Thresholded}& & \multicolumn{1}{c}{Unthresholded}& \\
    \multicolumn{2}{|c|}{} & FSL                                & AFNI       & SPM                     \\
    \hline
    {Preprocessing}        & {Motion Correction}                & \checkmark & \checkmark & \checkmark \\
    {}                     & {Segmentation}                     &            &            & \checkmark \\
    {}                     & {Brain Extraction (Anatomical)}    & \checkmark & \checkmark & \checkmark \\
    {}                     & {Brain Extraction (Functional)}    &            & \checkmark &            \\
    {}                     & {Intra-subject Coregistration}     & \checkmark & \checkmark & \checkmark \\
    {}                     & {Inter-subject Registration}       & \checkmark & \checkmark & \checkmark \\
    {}                     & {Analysis Voxel Size}              & \checkmark & \checkmark & \checkmark \\
    {}                     & {Smoothing}                        & \checkmark & \checkmark & \checkmark \\
    \hline
    {First-level}          & {Model Specification}              & \checkmark & \checkmark & \checkmark \\
    {}                     & {Inclusion of 6 Motion Parameters} & \checkmark & \checkmark & \checkmark \\
    {}                     & {Model Estimation}                 &            &            & \checkmark \\
    {}                     & {Contrasts}                        & \checkmark & \checkmark & \checkmark \\
    \hline
    {Second-level}         & {Model Specification}              & \checkmark & \checkmark & \checkmark \\
    {}                     & {Model Estimation}                 &            &            & \checkmark \\
    {}                     & {Contrasts}                        &            & \checkmark & \checkmark \\
    {}                     & {Second-level Inference}           & \checkmark & \checkmark & \checkmark \\
    \hline
  \end{tabular}
  \caption{Software processing steps (adapted from~\cite{bowring2019exploring}). \camille{I don't fully understand the "first-level" and "second-level" info in this table. Do to the "contrasts" the model has to be estimated so each time "Contrast" is ticked then "model estimation" should be ticked as well}}
  \label{table:pipeline-steps}
\end{table}

\subsection{Fuzzy Libmath environment}

To introduce numerical noise in the analyses, we used
Fuzzy Libmath~\cite{salari2021accurate}, a version of the GNU
mathematical library (libmath) instrumented with Monte-Carlo arithmetic.
Monte-Carlo arithmetic simulates numerical errors
by introducing a controlled amount of noise in floating-point
operations through the following perturbation~\cite{Parker1997-qq}:

\begin{equation} \label{eq:mca_inexact}
  inexact(x) = x + 2^{e_x-t}\xi,
\end{equation}
where $e_x$ is the exponent in the floating-point representation of $x$,
$t$ is the virtual precision (the number of unperturbed bits in the
mantissa of $x$), and $\xi$ is a random uniform variable of
$(-\frac{1}{2}, \frac{1}{2})$. We introduced the perturbation using
Verificarlo~\cite{denis2015verificarlo}, an LLVM compiler supporting Monte-Carlo
arithmetic and other types of numerical instrumentations.

We loaded the instrumented libmath functions in the pipeline using
LD\_PRELOAD, a Linux mechanism to force-load a shared library into an
executable. This mechanism allows functions defined in Fuzzy Libmath to transparently
overload the original ones without the need to modify or recompile the
analysis pipeline.

Fuzzy Libmath introduces numerical perturbations in the values returned by
mathematical functions but not in their input values or within their
implementation. This is done by wrapping the original functions and
applying function \texttt{inexact} to their returned values.
Listing~\ref{algo:wrapper} shows an example of this wrapping for the
\texttt{log} function in single and double precision. In this wrapper, the
original function is called through \texttt{dlsym}, a function that returns
the memory address of a symbol --- in our case \texttt{RTLD\_NEXT}, the
address of the next occurrence of the function in memory. Compiling function wrappers
with Verificarlo instruments the result of the
addition between the original function output and the floating-point zero.

\lstdefinestyle{customasm}{
  belowcaptionskip=1\baselineskip,
  frame=L,
  xleftmargin=\parindent,
  language=[x86masm]Assembler,
  basicstyle=\footnotesize\ttfamily,
  commentstyle=\itshape\color{purple!40!black},
}
\lstinputlisting[caption=Sample wrapper function (C code) \gk{I don't think we need this/it adds anything. Could go to a supplement if we want it in a paper somewhere}, label=algo:wrapper, style=customasm]{wrapper.c}

In~\cite{salari2021accurate}, Fuzzy Libmath was shown to accurately
simulate the effect of Linux operating system updates in structural
pre-processing pipelines of the Human Connectome Project which are largely based on FSL.
To validate our pipeline instrumentations for the present study, we first verified that non-instrumented
executions of the same pipeline on the same dataset led to identical
results. We also listed the pipeline
library dependencies using the \texttt{ldd} Linux utility and verified that
(1) the tested pipelines were dynamically linked to the GNU libmath library, and
(2) there was no alternative implementation of elementary mathematical functions in the pipeline dependencies.
Finally, we verified that the use of Fuzzy Libmath affected computational results.

\subsection{Data processing}

We measured between-tool (BT) variability by running the pipelines
described in~\cite{bowring2019exploring} with FSL version 5.0.10, AFNI
version 18.1.09, and SPM12 version r7771 executed with GNU/Octave version
5.2. These software versions were identical to the study
in~\cite{bowring2019exploring} except for SPM for which we used GNU/Octave
instead of MATLAB to enable mathematical function instrumentation using
Fuzzy Libmath. Indeed, MATLAB uses its own built-in mathematical functions,
which prevents the use of Fuzzy Libmath. In AFNI, we set the number of
threads to 7 even though AFNI executions
in~\cite{bowring2019exploring} were single-threaded. This was meant to
reduce the time overhead resulting from Fuzzy Libmath instrumentation
\gk{Did we verify that single-threading generally didn't lead to different results than 7-threading? As in, were they all ostensibly sampled from the same random distribution?}.
\ali{We verified that multithreading leads to different results (e.g. single-threading vs, 7-threading).
  However, single-threaded executions couldn't reproduce the original results.}
All the analyses were conducted on the CentOS 7.3 operating system. The
computations were performed on \href{https://www.computecanada.ca}{Compute
  Canada's} Béluga cluster nodes, each with 2$\times$ Intel Gold 6148 Skylake~@~2.4~GHz
(40 cores/node) CPU and 8~GB of RAM per core. To facilitate portability and reproducibility,
we encapsulated the
above-mentioned software packages in Docker container images based on CentOS 7.3.
Although, we converted Docker containers to Singularity images to enable running on cluster nodes.

We measured numerical (within-tool; WT) variability by running the same analyses three
times using Fuzzy Libmath with a virtual precision of $t=53$~bits for
double-precision values and $t=24$~bits for single-precision values. These
values were chosen such that the numerical perturbation simulates machine
error. The resulting samples were equally plausible estimates of
the true \yohan{How can you assert that when you don't have ground truth? Would it be more safe to say "plausible estimates numerical solutions"?} numerical result at the precision used by the pipelines. Moreover, to evaluate numerical perturbations at different magnitudes,
we repeated the FSL analyses for virtual
precisions ranging from $t=1$~bit to $t=24$~bits for single-precision
and double-precision values. We also repeated the FSL analyses for virtual
precisions ranging from $t=24$~bits to $t=53$~bits for double-precision values, having
set the virtual precision to $t=24$~bits for single-precision values.

We evaluated WT and BT variability for thresholded as well as unthresholded
group-level and subject-level t-statistics maps by computing the
standard deviation of t-statistic maps across pipelines or Fuzzy Libmath samples, respectively. For BT, we computed the standard deviation
across a given pair of tools \camille{standard dev with 2 samples ? --> cf. discussion on mattermost).
For WT, we computed the standard deviation across the three
Fuzzy Libmath samples of a given tool. We computed WT variability for a pair of tools (A, B) as follows:
\begin{equation}
  \sigma_{WT(A,B)}^2 = \sigma_{WT(A)}^2 + \sigma_{WT(B)}^2,
  \label{eq:wt-pair}
\end{equation}
where $\sigma_{WT(.)}$ is the numerical variability for a given tool
and $\sigma_{WT(., .)}$ is the numerical variability for a pair of tools.

Further, from the thresholded maps, we determined regional instability
between activation clusters in the 360 regions in the Human Connectome
Project Multi-Modal Parcellation atlas version 1.0
(HCP-MMP1.0)~\cite{glasser2016multi}. For BT, we considered a region
unstable for a pair of tools if it contained activated voxels for a tool
but not for the other one. For WT, we considered a region unstable for a
pair of tools (A, B) if for tool A or tool B it contained activated voxels only for some Fuzzy Libmath
samples. \yohan{A voxel is activated if its intensity is $>$ 0? If so, the given definition focuses on regions with low intensities while regions with high intensities may be more unstable}

\section{Results}
The scripts, Docker images, and data to reproduce the results are available
in our GitHub repository at
\url{https://github.com/big-data-lab-team/fuzzy-neurotools}.

\subsection{Validation of Replication}

We verified the correctness of our analyses by comparing our unperturbed
t-statistic group maps with the ones obtained
in~\cite{bowring2019exploring}. For FSL, we found
identical results. For SPM, the files (as measured by checksums) were different but differences
were visually unnoticeable. For AFNI, the activation maps were similar
overall, however, differences were noticeable visually.
The observed differences remained small (see Supplemental Material~\ref{sec:supp-repro}), and might be due to the use of
GNU/Octave vs MATLAB in SPM, and of multithreading in AFNI. We performed visual quality control of the AFNI
and SPM results for each individual subject and confirmed that T1-weighted images were
correctly skull-stripped and registered to the MNI template.

\subsection{In the group analysis, BT variability was larger than WT variability}

Table~\ref{table:pipeline-stats} presents summary statistics for the
group-level t-statistics. For each tool pair (A, B), BT variability was
significantly larger than machine error in tool A or B, in both thresholded
and unthresholded maps (Wilcoxon signed-rank test and t-test p~\textless~$10^{-5}$ for all tests).
% \gk{we don't have sufficient sample size to say that p is that small,
% so we should probably say less than $10^{-5}$ or something else we could actually
% arrive at with our number of repetitions, pipeline pairs, and samples.}
% for all tests, t-test p=0 \gk{similar comment re p value} for all tests) \camille{p is smaller than a very small p but not zero?}.
Figure~\ref{fig:unthresh-maps}-\textbf{A},\textbf{B},\textbf{C} shows that
these global differences were confirmed regionally.
In addition, BT and WT variability appeared moderately correlated across voxels (Pearson's R
in [0.56, 0.58], p=0.0, Figure~\ref{fig:unthresh-maps}-\textbf{D}). Interestingly, the correlation
appears driven by a set of voxels exhibiting comparable BT and
WT values located around the identity line in Figure~\ref{fig:unthresh-maps}D, which, however, did not seem to have
any spatial consistency.

\setlength{\tabcolsep}{5pt}
\begin{table}[h]
  \centering
  \begin{tabular}{cccccc|cc}
    \toprule
                      &              & \multicolumn{4}{c|}{Group map}  & \multicolumn{2}{c}{Subject maps}                                                                     \\
    \multirow{2}{*}{} & {}           & \multicolumn{2}{c}{Thresholded} & \multicolumn{2}{c|}{Unthresholded} & \multicolumn{2}{c}{Unthresholded}                               \\
    % \cmidrule{3-8}
    {}                & {}           & $\mu$                           & $\sigma$                           & $\mu$                             & $\sigma$ & $\mu$ & $\sigma$ \\
    \midrule
    \rowcolor{lightgray!50}
    {Between Tools}   & FSL vs. SPM  & 1.282                           & 0.525                              & 0.443                             & 0.344    & 0.366 & 0.293    \\
    \rowcolor{lightgray!50}
    {(BT)}            & FSL vs. AFNI & 1.548                           & 0.616                              & 0.547                             & 0.441    & 0.439 & 0.352    \\
    \rowcolor{lightgray!50}
    {}                & AFNI vs. SPM & 1.475                           & 0.672                              & 0.608                             & 0.477    & 0.491 & 0.381    \\
    {Within Tool}     & FSL          & 0.354                           & 0.491                              & 0.082                             & 0.065    & 0.077 & 0.054    \\
    {(WT)}            & SPM          & 0.252                           & 0.448                              & 0.054                             & 0.045    & 0.048 & 0.037    \\
    {}                & AFNI         & 0.434                           & 0.524                              & 0.128                             & 0.135    & 0.108 & 0.131    \\
    \bottomrule
  \end{tabular}
  \caption{Voxel-wise mean and standard deviation of BT and WT variability
    in t-statistics maps.}
  \label{table:pipeline-stats}
\end{table}

\begin{figure*}[ht]
  \centering
  % \fbox{\begin{minipage}{\dimexpr \textwidth-2\fboxsep-2\fboxrule}
  \begin{subfigure}[ht]{0.9\textwidth}
    \centering
    \includegraphics[width=0.9\textwidth]{figures/std/gl-unthresh.png}
    %\caption{Standard deviation of thresholded t-statistics map on template surface}
  \end{subfigure}
  \caption{Unthresholded group-level variability computed between tools
    (\textbf{A}), within tools at machine error (\textbf{B}), difference
    between them (\textbf{C}), and voxel-wise comparison (\textbf{D}). }
  \label{fig:unthresh-maps}
  % \end{minipage}}
\end{figure*}

\subsection{In subject analyses, WT variability approached BT variability in some regions}

Table~\ref{table:pipeline-stats} also presents summary statistics for
subject-level unthresholded t-statistics maps. As for group-level maps,
BT variability was consistently larger than WT variability (Wilcoxon
signed-rank test and t-test p~\textless~$10^{-5}$ for all tests).
However, Figure~\ref{fig:unthresh-maps-sbj} shows that for
some subjects, WT variability approached and even
surpassed BT in some regions. As for the group maps, BT and WT
variability appeared moderately correlated for all subjects (R in [0.442,
    0.612], p=0.0 for all subjects, see Supplemental Material~\ref{sec:supp-subjects}).

%%%%%%%%%% Var. of Unthresh sbj05%%%%%%%%
\begin{figure*}[ht]
  \centering
  \includegraphics[width=0.9\textwidth]{figures/std/sbj05-std.png}
  %\caption{Standard deviation of thresholded t-statistics map on template surface}
  \caption{For subject with highest WT variability,
    unthresholded subject-level variability computed between tools (\textbf{A}), within tools at machine error (\textbf{B}),
    and difference between them (\textbf{C}). \camille{If the changes are visible, I think that it would be useful to add the correspondant unthresholded maps for the reader to see where the differences are and what they "look like"}}
  \label{fig:unthresh-maps-sbj}
\end{figure*}

\subsection{At precision t=17~bits, WT variability approached BT variability in the group analysis}

While the previous results were obtained for machine error, we also
evaluated numerical variability across different virtual precisions for FSL
and found that the virtual precision of t=17 bits minimized the RMSE
between BT and WT in unthresholded group maps.
FSL produced group maps with $\mu=0.325$ and $\sigma=0.287$ at this virtual precision,
which indicates a bit lower variability compared to BT variability
(Wilcoxon signed-rank test and t-test p~\textless~$10^{-5}$ for all tests).
Figure~\ref{fig:gnp-mni} shows BT
and WT in the corresponding group maps obtained at t=17 bits, showing that
BT and WT reach comparable magnitudes in some regions. BT and WT remained
moderately correlated at this precision.

%%%%%%%%%% plot different precisions%%%%%%%%
\begin{figure*}[ht]
  \begin{subfigure}[ht]{\textwidth}
    \centering
    \includegraphics[width=.75\textwidth]{figures/bg_global_precision.pdf}
    %\caption{Standard deviation of thresholded t-statistics map on template surface}
  \end{subfigure}
  %\caption{Standard deviation of thresholded t-statistics map on template surface}
  \caption{Unthresholded group t-statistics standard deviations computed between tools (\textbf{A}),
    within tools at the virtual precision of t=17~bits (\textbf{B}), difference between them (\textbf{C}), and
    voxel-wise comparison (\textbf{D}).}
  \label{fig:gnp-mni}
\end{figure*}

\subsection{Previous results were confirmed in thresholded group maps}

Figure~\ref{fig:thresh-maps}-\textbf{A},\textbf{B},\textbf{C} compares BT
and WT for thresholded group maps. Thresholding is an unstable operation
that introduced variability at the edges of
active regions for both BT and WT. Except in these regions, BT remained consistently larger
than WT \gk{what regions? This kind of came from nowhere}. Moreover, to measure correlation between BT and WT, we
measured WT instability and BT instability in each region of the
HCP-MMP1.0 parcellation \gk{again, it feels kind of abrupt and I'm not really sure what this section is getting at or how it differs from the previous based on these descriptions. Perhaps introduce that thresholding is common, but since it makes things discontinuous we study it through aparcellation, etc...}. The confusion matrices in
Figure~\ref{fig:thresh-maps}-\textbf{D} report these instabilities for
the 360 tested regions. The average ratio of unstable regions was 26\%
for BT and 9.2\% for WT, which confirmed that BT variability was larger than machine error.
The average Cohen's kappa score\footnote{$\kappa \leq 0$ denotes chance agreement, $-1 \leq \kappa \leq 1$}
between WT instability and BT instability was $\kappa=0.17$, indicating
a moderate agreement between WT instability and BT instability.

%%%%%%%%%% Var. of Thresh %%%%%%%%
\begin{figure*}[ht]
  \begin{subfigure}[ht]{\textwidth}
    \centering
    \includegraphics[width=\textwidth]{figures/std/gl-thresh.png}
    %\caption{Standard deviation of thresholded t-statistics map on template surface}
  \end{subfigure}
  \centering
  \caption{Thresholded group t-statistics standard deviations computed between tools (\textbf{A}),
    within tools at the virtual precision of t=17~bits (\textbf{B}), difference between them (\textbf{C}), and
    confusion matrices of activation instability
    in BT and WT among the 360 regions of the HCP-MMP1.0 parcellation (\textbf{D}).}
  \label{fig:thresh-maps}
\end{figure*}

\section{Discussion}

In fMRI group analyses, within-tool variability remains an order of magnitude smaller
than between-tool variability. Group analyses benefit from regularization
of numerical noise, which is expected to increase as sample size
increases. Therefore, for fMRI studies with large sample sizes, within-tool variability may be negligible with respect to between-tool variability. The
recommendation made in~\cite{botvinik2020variability} to rely on
``multiverse'' analyses where multiple analysis tools are compared is
therefore likely to successfully correct for machine error in group
studies. Nevertheless, within-tool variability remains substantial in group analyses
that are based on a single tool, as is commonly the case in current fMRI
studies. In particular, in our study, the inherent instability of
thresholding was triggered by machine error in 9\% of 360 brain regions,
which indicates that machine error might have impacted neuroscientific
conclusions related to these regions.

In subject-level analyses, within-tool and between-tool variabilty can become
of comparable magnitude for some subjects in some regions. This observation
is particularly relevant to the development of fMRI-based biomarkers aiming
at individualized phenotype predictions. Machine error is expected to play
a non-negligible role in such analyses, even when predictions combine
results produced by multiple tools. The observed regularization effect of
group analyses toward numerical noise is consistent with observations made
in~\cite{kiar2020numerical} from diffusion MRI where connectome graph
statistics were found to be substantially unstable at the subject-level
while group distributions remained consistent. \gk{this last sentence feels like it belongs in the previous paragraph, right before mentioning thresholding}

For both group- and subject-level analyses, between-tool variability and within-tool
variability were found to be moderately correlated. Even though between-tool
variability and within-tool  variability are different in nature, this result
suggests that in some cases both sources of variability may have a common
cause that might be related to the conditioning of the BOLD signal
estimation problem \gk{I wouldn't necessarily say it's because of "BOLD signal estimation problem", just "algorithms used to process"} in a specific dataset. Therefore, in some regions,
instabilities of similar magnitude may be triggered by small numerical
perturbations, model variations, or implementation differences. For
instance, the analysis of datasets with high motion may be unstable both
across tools and numerically.

Therefore, numerical stability may be a suitable proxy to study
between-tool variability. This speculation might be of practical value to
address software variability at large given that numerical stability refers
to a consistent mathematical framework whereas between-tool variability
remains more empirical. Likewise, useful quality control metrics may be derived from numerical and between-tool stability.

The finding that numerical variability approached tool variability at the
virtual precision of t=17~bits is interesting too. Indeed, while machine
error generally introduces differences in the order of \yohanmod{1} \yohan{1/2} ulp
\yohan{Are we talking about one arithmetic operation? Or an entire execution?}
--- or
t=53~bits for double-precision values and t=24~bits for single-precision
ones \yohan{I would not mix virtual precision and error measurement}--- common scientific software dependencies introduce larger errors
.
For instance, SciPy's 2D spline interpolation was recently found to be
precise up to t=10~bits~\cite{pytracer} \tristan{Fix ref}
\yohan{Same as previous, I would talk about significance or relative error, or remove the t= and not talk about virtual precision} and might therefore introduce
numerical perturbations leading to errors in the range of between-tool
variability \tristan{Yohan, please check that sentence}. Such high errors
might be triggered by updates in operating systems, Python, MATLAB, and
other software dependencies.
\yohan{The sentence is ok, just a picky point: the SciPy result you cited describes the precision of the spline implementation while at the end you're talking differences due to updates.
  The nuance is that two versions may have a similar precision and so when you udpate your
  system, the difference between both versions is low but each version can be ill-conditionned.}

Our results are limited by the type of numerical noise introduced in the
analyses. We only perturbed the outputs of elementary
mathematical functions while numerical noise could creep in any
floating-point operation. Our estimation of numerical variability should
therefore be considered a lower bound. Our estimation of tool varability is
also likely to be underestimated, having tested only 3 analytical pipelines
among the thousands available~\cite{carp2012plurality}.

In conclusion, our results motivate further numerical stability
investigations in fMRI analysis pipelines. Pipeline-level analyses could be
conducted to identify specific components that contribute to numerical
variability, and if possible correct them accordingly. When instability is
inherent to the analysis, sampling results distributions through numerical
perturbations might improve stability, as explored in~\cite{kiar2021data}.
Finally, pipeline-specific statistical corrections might be envisaged to
account for numerical variability.

\bibliographystyle{plain}
\bibliography{biblio}

\clearpage

\setcounter{equation}{0}
\setcounter{figure}{0}
\setcounter{table}{0}
\setcounter{section}{0}

\makeatletter
\renewcommand{\theequation}{S\arabic{equation}}
\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\thesection}{S\arabic{section}}

\textbf{\centering \Large Supplemental Materials}

\section{Reproduced results}
\label{sec:supp-repro}

Figure~\ref{fig:replication-diff} shows the difference in SPM and AFNI
group analyses maps between the results in~\cite{bowring2019exploring} and
our replication. Numerical perturbations of 1 ulp are likely to have been
introduced by our replication due to the use of GNU/Octave vs MATLAB for SPM
and multithreading for AFNI \camille{Would this be feasible to check?}. However, the group maps remained very similar
overall, which led us to conclude that our results correctly reproduced
the ones in~\cite{bowring2019exploring}.
\begin{figure*}[ht]
  % \fbox{\begin{minipage}{\dimexpr \textwidth-2\fboxsep-2\fboxrule}
  % \centering
  \includegraphics[width=\textwidth]{figures/replication-diffs.png}
  \caption{Differences between reproduced and original results obtained in~\cite{bowring2019exploring}
    of unthresholded group-level t-statistics for SPM (left) and AFNI (right). \camille{it looks like the highest areas of difference in AFNI are due to difference un masking maybe worth noting?}}
  \label{fig:replication-diff}
  % \end{minipage}}
\end{figure*}

\section{BT and WT correlations for all subjects}
\label{sec:supp-subjects}

Figure~\ref{fig:unthresh-correlation-allsbj} plots the relationship between
WT and BT for each subject, showing a consistent moderate correlation between BT and
WT.

\begin{figure*}[ht]
  \centering
  \includegraphics[width=.6\textwidth]{figures/sbj-std-corr-unthresh-plot.png}
  \caption{Comparison between BT and WT variabilty for 16 subjects.}
  \label{fig:unthresh-correlation-allsbj}
\end{figure*}

\end{document}
\endinput

### 2021-12-16 paper/paper.tex
Update variability computations with absolute diffs

@@
     {}                     & {Inclusion of 6 Motion Parameters} & \checkmark & \checkmark & \checkmark \\
-    {}                     & {Model Estimation}                 &            &            & \checkmark \\
+    {}                     & {Model Estimation}                 & \checkmark & \checkmark & \checkmark \\
     {}                     & {Contrasts}                        & \checkmark & \checkmark & \checkmark \\
@@
     {Second-level}         & {Model Specification}              & \checkmark & \checkmark & \checkmark \\
-    {}                     & {Model Estimation}                 &            &            & \checkmark \\
-    {}                     & {Contrasts}                        &            & \checkmark & \checkmark \\
+    {}                     & {Model Estimation}                 & \checkmark & \checkmark & \checkmark \\
+    {}                     & {Contrasts}                        & \checkmark & \checkmark & \checkmark \\
     {}                     & {Second-level Inference}           & \checkmark & \checkmark & \checkmark \\
@@
 \camille{I don't fully understand the "first-level" and "second-level" info in this table.
-Do to the "contrasts" the model has to be estimated so each time "Contrast" is ticked then "model estimation" should be ticked as well}}
+Do to the "contrasts" the model has to be estimated so each time "Contrast" is ticked then "model estimation" should be ticked as well}\ali{fixed.}}
   \label{table:pipeline-steps}
@@
 group-level and subject-level t-statistics maps by computing the
-standard deviation of t-statistic maps across pipelines or Fuzzy Libmath samples, respectively. For BT, we computed the standard deviation
-across a given pair of tools \camille{standard dev with 2 samples ? --> cf. discussion on mattermost).
-For WT, we computed the standard deviation across the three
+absolute differences of t-statistic maps across pipelines or Fuzzy Libmath samples, respectively.
+For BT, we computed the absolute differences across a given pair of tools.
+For WT, we averaged the absolute differences between three pair combinations across the three
 Fuzzy Libmath samples of a given tool. We computed WT variability for a pair of tools (A, B) as follows:
 \begin{equation}
-  \sigma_{WT(A,B)}^2 = \sigma_{WT(A)}^2 + \sigma_{WT(B)}^2,
+  WT(A,B) = \frac{WT(A) + WT(B)}{2},
   \label{eq:wt-pair}
 \end{equation}
-where $\sigma_{WT(.)}$ is the numerical variability for a given tool
-and $\sigma_{WT(., .)}$ is the numerical variability for a pair of tools.
+where $WT(.)$ is the numerical variability for a given tool
+and $WT(., .)$ is the numerical variability for a pair of tools.
 
@@
 In addition, BT and WT variability appeared moderately correlated across voxels (Pearson's R
-in [0.56, 0.58], p=0.0, Figure~\ref{fig:unthresh-maps}-\textbf{D}). Interestingly, the correlation
-appears driven by a set of voxels exhibiting comparable BT and
-WT values located around the identity line in Figure~\ref{fig:unthresh-maps}D, which, however, did not seem to have
-any spatial consistency.
+in [0.60, 0.62], p=0.0, Figure~\ref{fig:unthresh-maps}-\textbf{D}). 
+% Interestingly, the correlation appears driven by a set of voxels exhibiting comparable BT and
+% WT values located around the identity line in Figure~\ref{fig:unthresh-maps}D, which, however, did not seem to have
+% any spatial consistency.
 
@@
     \rowcolor{lightgray!50}
-    {Between Tools}   & FSL vs. SPM  & 1.282                           & 0.525                              & 0.443                             & 0.344    & 0.366 & 0.293    \\
+    {Between Tools}   & FSL vs. SPM  & 2.564                           & 1.050                              & 0.887                             & 0.689    & 0.733 & 0.586    \\
     \rowcolor{lightgray!50}
-    {(BT)}            & FSL vs. AFNI & 1.548                           & 0.616                              & 0.547                             & 0.441    & 0.439 & 0.352    \\
+    {(BT)}            & FSL vs. AFNI & 3.096                           & 1.232                              & 1.095                             & 0.882    & 0.878 & 0.704    \\
     \rowcolor{lightgray!50}
-    {}                & AFNI vs. SPM & 1.475                           & 0.672                              & 0.608                             & 0.477    & 0.491 & 0.381    \\
-    {Within Tool}     & FSL          & 0.354                           & 0.491                              & 0.082                             & 0.065    & 0.077 & 0.054    \\
-    {(WT)}            & SPM          & 0.252                           & 0.448                              & 0.054                             & 0.045    & 0.048 & 0.037    \\
-    {}                & AFNI         & 0.434                           & 0.524                              & 0.128                             & 0.135    & 0.108 & 0.131    \\
+    {}                & AFNI vs. SPM & 2.951                           & 1.344                              & 1.218                             & 0.955    & 0.983 & 0.762    \\
+    {Within Tool}     & FSL          & 0.510                           & 0.697                              & 0.127                             & 0.098    & 0.120 & 0.083    \\
+    {(WT)}            & SPM          & 0.367                           & 0.638                              & 0.085                             & 0.068    & 0.075 & 0.056    \\
+    {}                & AFNI         & 0.637                           & 0.747                              & 0.199                             & 0.202    & 0.167 & 0.197    \\
     \bottomrule
@@
     \centering
-    \includegraphics[width=0.9\textwidth]{figures/std/gl-unthresh.png}
+    \includegraphics[width=0.9\textwidth]{figures/abs/gl-unthresh.png}
     %\caption{Standard deviation of thresholded t-statistics map on template surface}
@@
 However, Figure~\ref{fig:unthresh-maps-sbj} shows that for
-some subjects, WT variability approached and even
-surpassed BT in some regions. As for the group maps, BT and WT
-variability appeared moderately correlated for all subjects (R in [0.442,
-    0.612], p=0.0 for all subjects, see Supplemental Material~\ref{sec:supp-subjects}).
+some subjects, WT variability approached %and even surpassed
+BT in some regions. As for the group maps, BT and WT
+variability appeared moderately correlated for all subjects (R in [0.484,
+    0.633], p=0.0 for all subjects, see Supplemental Material~\ref{sec:supp-subjects}).
 
@@
   \centering
-  \includegraphics[width=0.9\textwidth]{figures/std/sbj05-std.png}
+  \includegraphics[width=0.9\textwidth]{figures/abs/sbj05-abs.png}
   %\caption{Standard deviation of thresholded t-statistics map on template surface}
@@
 between BT and WT in unthresholded group maps.
-FSL produced group maps with $\mu=0.325$ and $\sigma=0.287$ at this virtual precision,
+FSL produced group maps with $\mu=0.483$ and $\sigma=0.410$ at this virtual precision,
 which indicates a bit lower variability compared to BT variability
@@
   %\caption{Standard deviation of thresholded t-statistics map on template surface}
-  \caption{Unthresholded group t-statistics standard deviations computed between tools (\textbf{A}),
+  \caption{Unthresholded group t-statistics absolute differences computed between tools (\textbf{A}),
     within tools at the virtual precision of t=17~bits (\textbf{B}), difference between them (\textbf{C}), and
@@
     \centering
-    \includegraphics[width=\textwidth]{figures/std/gl-thresh.png}
+    \includegraphics[width=\textwidth]{figures/abs/gl-thresh.png}
     %\caption{Standard deviation of thresholded t-statistics map on template surface}
@@
   \centering
-  \caption{Thresholded group t-statistics standard deviations computed between tools (\textbf{A}),
-    within tools at the virtual precision of t=17~bits (\textbf{B}), difference between them (\textbf{C}), and
+  \caption{Thresholded group-level variability computed between tools (\textbf{A}),
+    within tools at machine error (\textbf{B}), difference between them (\textbf{C}), and
     confusion matrices of activation instability
@@
   \centering
-  \includegraphics[width=.6\textwidth]{figures/sbj-std-corr-unthresh-plot.png}
+  \includegraphics[width=.6\textwidth]{figures/sbj-abs-corr-unthresh-plot.png}
   \caption{Comparison between BT and WT variabilty for 16 subjects.}

### 2021-12-16 computs.py

@@ -14,6 +14,7 @@ from statistics import mean
 import scipy.stats as stat
 from scipy.signal import resample
 import os
+import pandas as pd
 
 
 ############# Global nearest precision
@@ -38,27 +39,24 @@ def global_nearest_precision():
     rmse_ = {}
     all_rmse = {}
     global_average_rmse = []
-
     for p in range(11, 54, 2):
         rmse_[p] = 0
         all_rmse[str(p)] = []
         for s in ['fsl-afni', 'fsl-spm', 'afni-spm']:
-            bt_ = nib.load('./data/std/{}-unthresh-std.nii.gz'.format(s))
+            bt_ = nib.load('./data/abs/{}-unthresh-abs.nii.gz'.format(s))
             # bt_ = bt_.get_fdata()
-            wt_fsl = nib.load('./data/std/FL-FSL/p{}_fsl_unthresh_std.nii.gz'.format(p))
+            wt_fsl = nib.load('./data/abs/FL-FSL/p{}_fsl_unthresh_abs.nii.gz'.format(p))
             # wt_fsl = wt_fsl.get_fdata()
             rmse_value = calculate_rmse(bt_, wt_fsl)
             all_rmse[str(p)].append(rmse_value)
             rmse_[p] += rmse_value
 
     all_rmse = np.transpose(np.array(list(all_rmse.values())))
-
     return min(rmse_, key=rmse_.get), all_rmse
 
 
 def plot_rmse_nearest(all_rmse):
     import matplotlib.cm as cm
-
     fig = plt.figure(figsize=(15,10))
     colors = ['red', 'green', 'blue']
     tool_pair = ['fsl-afni', 'fsl-spm', 'afni-spm']
@@ -66,30 +64,27 @@ def plot_rmse_nearest(all_rmse):
         #type_ = ""
         #if s+1 in i2T1w: type_ += "*"
         #if s+1 in i2T2w: type_ += "$\dag$"
-        plt.plot(range(11,54, 2), rmse_, marker='x', color=colors[s], alpha=0.8, label=tool_pair[s].upper())
+        plt.plot(range(11,54, 2), rmse_, marker='x', color=colors[s], alpha=0.8, label="({})".format(tool_pair[s].replace('-', ', ').upper()))
 
     average = np.mean(all_rmse, axis=0, dtype=np.float64)
[…]

### 2021-12-17 paper/paper.tex
Rewrote intro

--- 2021-12-13 paper/paper.tex
@@
   the relationship between numerical (within-tool) and between-tool variabilities for
-  given analysis is unclear. We extended a previous comparison
+  a given analysis is unclear. We extended a previous comparison
   of fMRI analysis software libraries (namely FSL, AFNI, and SPM) and relate the observed differences to the numerical
@@
   Finally, we found that numerical variability had effects of similar magnitude to between-tool
-  variability when errors were introduced at the precision of 17 bits --- a
-  precision level found in commonly-used libraries \gk{What does this mean?
-They only have precision of 17, or they use data that is 17 bits?}.
-\yohan{agree with Greg, it is unclear and a strong assertion!} \camille{It is not entirely clear to me how this relates to the previous conclusions that BT variability was consistently larger than machine error} Our findings motivate the
+  variability when numerical errors are introduced at the precision of 17 bits --- a
+  precision level found in commonly-used libraries \gk{What does this mean? They only have precision of 17, or they use data that is 17 bits?}.
+   \yohan{agree with Greg, it is unclear and a strong assertion!}
+   \camille{It is not entirely clear to me how this relates to the previous conclusions that BT variability 
+   was consistently larger than machine error} Our findings motivate the
   continued investigation of numerical instability in neuroimaging, and position it
@@
 
-Software variability has been highlighted as an important source of error
-in fMRI analysis, originating from mainly two factors: between-tool
-variability denotes differences in the outcomes of different
-software toolboxes --- such as FSL, AFNI and SPM --- when analyzing the same dataset
-using similar methods \gk{this sentence says "two factors" but then only mentioned between tool.
-Also, it is too long}.
-\yohan{What do you mean by methods?
-Numerical schemes?
-If so, are we sure that theses tools are using the same numerical schemes?} \camille{I also had highlighed the 2 here, I think we could use "multiple factors" --> in practice we could also add different types of within-software variability (in the form of changing algorithms and/or parameters for a given step)} Numerical variability, in contrast, denotes
-differences in outcomes of the same analytical toolbox
-executed in different environments – such as different operating systems or
-different hardware platforms \gk{you mention these two examples of numerical variability, but you use neither in this paper.
-Why not mention the one you use?}.
-This manuscript investigates the extent to
-which both types of variability relate to each other.
+Recent explorations of the analytical flexibility in brain imaging across
+tools, platforms, or teams, have demonstrated unexpected variability in
+results, even when analyzing identical data~\cite{botvinik2020variability}. Possible explanations for
+such discrepancies include methodological flexibility~\cite{carp2012plurality} and software
+variability, the focus of this paper. A recommendation to address this
+variability is to adopt a ``multiverse" approach that analyzes the same
+dataset multiple times in different software environments, and ultimately
+conclude from the resulting set of outcomes. However, the range of
+analytical conditions to be included in such multiverse analyses remains
+poorly defined, both because of the boundless set of tools and
+configurations, and that the precise causes of software variability remain
+unclear. This lack of clarity is in part because fMRI analyses depend on
+complex software stacks that leverage low-level libraries provided by the
+operating system (e.g.,mathematical functions), general scientific
+computing methods (e.g., optimization toolboxes), and specific fMRI
+analysis tools (e.g., spatial image normalization methods). At each level,
+conceptual and implementation differences across experiments may each
+create substantial variability in the analysis outcomes.
 
-Numerical variability stems from errors introduced in floating point
-computations by \yohanmod{numerical} rounding, cancellation, and absorption~\cite{muller2018handbook} \gk{only? What about poor conditioning + random effects + noise in data?}. In
-practice, the occurrence of such errors is affected by variations in
-hardware, software dependencies, compilation, parallelization, and most
-likely other factors \gk{what does "most likely other factors" mean?}. In the best case, the impact of such errors and the induced
-variations is in the range of \yohanmod{1} \yohan{1/2 ulp for round-to-nearest} unit in the last place (ulp) --- also known
-as machine error, the difference between two successive floating-point
-numbers. However, when computations depend on imprecise libraries,
-numerical error may vary by several ulps. For instance, the \yohan{maximal} numerical error
-of the complementary error function (erfc) is larger than 3 ulps in several
-mathematical libraries~\cite{zimmermann:hal-03141101}, implying that an update of this function could introduce
-numerical \yohanmod{perturbations} \yohan{differences} larger than 3 ulps. Due to numerical instability,
-machine error may amplify during its propagation through analyses \yohan{do you mean calculations?},
-resulting in considerable loss of precision in final results. In fMRI analysis,
-numerical variability produced by Linux updates has been shown to result in Dice
-coefficients ranging from 0.0 to 1.0 for ICA components produced by FSL
-MELODIC~\cite{Glatard2015} \gk{there isn't really context for what this result means, yet.
-Is this a lot?
-Dice hasn't been explained.
-Also, how are ICA components (weighted vectors) compared via Dice (a measure for binary image/vector comparison)?
-}.
+As a result of these factors, the variability resulting from the use of
+different fMRI analysis tools implementing similar analytical approaches
+(between-tool variability) can reach worrying magnitudes. For instance, the
+study in~\cite{bowring2019exploring} compared the results produced by the
+three main fMRI analysis toolboxes, namely SPM, AFNI and FSL, using similar
+pipelines. It reported limited similarity between the activation clusters
+produced by these tools, measured by Dice coefficients ranging from 0.0 to
+0.684 \tristan{check values in erratum}, where 0 indicates non-overlapping
+clusters and 1 means identical clusters. More recently, the work
+in~\cite{Li2021.12.01.470790} also showed a low similarity between the
+results produced by five different tools (ABCD, CCS, CPAC, DPARSF, fMRIPrep
+\tristan{Ali, could you add refs to these tools from Li et al 2021? Also
+for SPM, FSL and AFNI in the previous sentence}) and identified the main
+factors contributing to these differences. The magnitude of the highlighted
+differences suggest that between-tool variability may be playing a critical
+role in the reproducibility of fMRI analysis overall.
 
-Between-tool variability is not a surprising concept: different programs
-executed on the same data may obviously produce different results \gk{even when doing conceptually the same thing? Why? Why is this obvious?}. However,
-in fMRI analysis, programs all aim to estimate brain activity from measured
-variations in the BOLD signal, which is not expected to be measurably
-impacted by computational artifacts \gk{Why wouldn't it be?
-Especially if you state that it isn't surprising for tools to lead to different results, how come you expect fMRI to not suffer from this?}.
-\yohan{+1} Therefore, the observation reported
-in~\cite{bowring2019exploring} that the three main fMRI analysis toolboxes
-may result in marked analytical differences (Dice coefficients ranging from
-0.000 to 0.684 \camille{Apologies I don't have the numbers on the top of my head but can I confirm taht those were taken from the erratum that was published after the paper?} for thresholded activation maps), has raised substantial
-concern. In a follow-up study~\cite{bowring2021isolating}, the same authors
-isolated the analytical components where the pipelines diverge,
-highlighting a major impact of the signal model and suggesting that the
-root-cause for between-tool variability depends on the analysis design and
-task paradigm \gk{If you state this finding, you need to unpack what it means a little bit.
-What part of analysis design?
-By task paradigm do you essentially mean the expected effect-size for the task that individuals are performing within the scanner?} \camille{would be worth adding here that the identified main sources of variations were different across the 3 studies}.
-Further, the work in~\cite{Li2021.12.01.470790} found
-moderate agreement between results produced by five independent fMRI
-analysis pipelines and analyzed in details the components that led to these
-variations \gk{This sentence feels inconsistent and counter to the previous, and also doesn't tell us their findings; what components led to variations?}.
-Between-tool variability is supposedly an important determinant
-of the substantial analytical variability recently observed in
-fMRI~\cite{botvinik2020variability} \gk{arguably, one of the larger issues that the NARPS result showed us was that people construct and test ``equivalent'' hypotheses differently, so despite highly similar data they will arrive at different conclusions}.
+The variability resulting from differences in lower-level software
+libraries (within-tool variability) has also been quantified in fMRI. The
+study in~\cite{Glatard2015} mentions a low similarity between the activation clusters
+produced by FSL using different versions of the GNU/Linux system, measured
+by Dice coefficients ranging from 0.0 to 1.0, covering the full spectrum of
+possible similarities. This variability resulted from updates in the GNU
+mathematical library and can be properly simulated by introducing small
+numerical perturbations on the results returned by mathematical functions~\cite{salari2021accurate}.
+Here again, these observations could partly explain the differences
+reported in~\cite{botvinik2020variability}.
 
-We investigate the relationship between numerical and between-tool
-variability through two main questions: (1) what is the relative magnitude of
-within-tool (numerical) variability and between-tool variability, and (2) is there an association between within- and between-tool variability.
-The first question determines the extent to which each variability
-source may impact current fMRI results, and which one should be addressed or leveraged
-in priority. The second question may provide deeper insights on the origins
-of software variability. Conceptually, both types of variability may
-reflect a single solution space shaped by local minima and
-ill-conditioning \gk{this feels a bit clumsily worded, but I don't immediately have a rewrite that comes to mind }, which could be explored through numerical or software
-perturbations.
+The relationship between within- and between-tool variability are poorly
+understood, but could play an important role in the construction of
+multiverse study environments. Importantly, if associations exist between
+these two types of variability, they may both originate to some degree from
+the inherent instability of brain activity estimation from BOLD signal
+variations, and shed light on the numerical confidence of results. Under
+this hypothesis, small perturbations introduced by low-level software
+updates could trigger effects correlated with those created by tool
+variations.
+
+This paper investigates the relationship between numerical stability -- used
+as a proxy for within-tool variability -- and between-tool variability
+through two main questions:
+\begin{enumerate}
+\item what is the relative magnitude of within-tool variability and between-tool variability, and
+\item is there an association between within- and between-tool variability.
+\end{enumerate}
+
+We address these two questions by reproducing the study in~\cite{bowring2019exploring} and
+extending it with the addition of numerical perturbations of controlled
+magnitude.
 
@@
 standard deviation of t-statistic maps across pipelines or Fuzzy Libmath samples, respectively. For BT, we computed the standard deviation
-across a given pair of tools \camille{standard dev with 2 samples ? --> cf. discussion on mattermost).
+across a given pair of tools \camille{standard dev with 2 samples ? cf. discussion on mattermost}.
 For WT, we computed the standard deviation across the three

### 2021-12-17 paper/paper.tex

@@
 group-level and subject-level t-statistics maps by computing the
-standard deviation of t-statistic maps across pipelines or Fuzzy Libmath samples, respectively. For BT, we computed the standard deviation
-across a given pair of tools \camille{standard dev with 2 samples ? cf. discussion on mattermost}.
-For WT, we computed the standard deviation across the three
+absolute differences of t-statistic maps across pipelines or Fuzzy Libmath samples, respectively.
+For BT, we computed the absolute differences across a given pair of tools.
+For WT, we averaged the absolute differences between three pair combinations across the three
 Fuzzy Libmath samples of a given tool. We computed WT variability for a pair of tools (A, B) as follows:
 \begin{equation}
-  \sigma_{WT(A,B)}^2 = \sigma_{WT(A)}^2 + \sigma_{WT(B)}^2,
+  WT(A,B) = \frac{WT(A) + WT(B)}{2},
   \label{eq:wt-pair}
 \end{equation}
-where $\sigma_{WT(.)}$ is the numerical variability for a given tool
-and $\sigma_{WT(., .)}$ is the numerical variability for a pair of tools.
+where $WT(.)$ is the numerical variability for a given tool
+and $WT(., .)$ is the numerical variability for a pair of tools.
 

### 2021-12-17 paper/paper.tex
Updates in method and results

@@
 
-\title{Comparing variability across fMRI analysis software packages and numerical errors}
+\title{Comparing variability across and within fMRI analysis software packages}
 
@@
 \begin{abstract}
-  Variability has been broadly observed in functional MRI analyses across both
-  differences in analytic tools and numerical instabilities. However,
-  the relationship between numerical (within-tool) and between-tool variabilities for
-  a given analysis is unclear. We extended a previous comparison
-  of fMRI analysis software libraries (namely FSL, AFNI, and SPM) and relate the observed differences to the numerical
-  variability observed using Monte-Carlo arithmetic. We found that
-  between-tool variability was consistently larger than numerical variability in group-level analyses,
-  whereas numerical variability approached between-tool variability in some subject-level analyses.
-  Interestingly, between-tool variability and
-  machine error appeared moderately correlated ($r > 0.5$) in all tested conditions.
-  Finally, we found that numerical variability had effects of similar magnitude to between-tool
-  variability when numerical errors are introduced at the precision of 17 bits --- a
-  precision level found in commonly-used libraries \gk{What does this mean? They only have precision of 17, or they use data that is 17 bits?}.
+  Variability has been broadly observed in functional MRI analyses as a result of differences
+  both between and within analytic tools. However, the
+  relationship between within-tool and between-tool variabilities for a
+  given analysis is unclear. We extended a previous comparison of fMRI
+  analysis software libraries (namely FSL, AFNI, and SPM) and related the
+  observed differences to within-tool variability
+  simulated using Monte-Carlo arithmetic. In group analyses, we found that
+  between-tool variability was consistently larger than within-tool
+  variability simulated from numerical perturbations of the magnitude of
+  machine error. In subject analyses, the same within-tool variability
+  approached between-tool variability in some regions for some subjects.
+  Interestingly, within- and between-tool variabilities appeared moderately
+  correlated ($r > 0.5$) in all tested conditions. Finally, we found that
+  within-tool variability had effects of similar magnitude to between-tool
+  variability when simulated from numerical perturbations introduced at the virtual
+  precision of 17 bits --- a high but plausible level of uncertainty for
+  scientific computing libraries \gk{What does this mean? They only have
+  precision of 17, or they use data that is 17 bits?}.
    \yohan{agree with Greg, it is unclear and a strong assertion!}
-   \camille{It is not entirely clear to me how this relates to the previous conclusions that BT variability 
-   was consistently larger than machine error} Our findings motivate the
-  continued investigation of numerical instability in neuroimaging, and position it
+   \camille{It is not entirely clear to me how this relates to the previous
+   conclusions that BT variability was consistently larger than machine
+   error} \tristan{Better?} Our findings motivate the
+  continued investigation of within-tool variability in neuroimaging, and position it
   as a possible proxy for studying arbitrary between-tool variations.
@@
 produced by these tools, measured by Dice coefficients ranging from 0.0 to
-0.684 \tristan{check values in erratum}, where 0 indicates non-overlapping
+0.769, where 0 indicates non-overlapping
 clusters and 1 means identical clusters. More recently, the work
@@
 effect of tool variability for this dataset, which we intended to
-extend with the present quantification of numerical variability.
+extend with the present quantification of within-tool variability.
 
@@
   \end{tabular}
-  \caption{Software processing steps (adapted from~\cite{bowring2019exploring}).
-\camille{I don't fully understand the "first-level" and "second-level" info in this table.
-Do to the "contrasts" the model has to be estimated so each time "Contrast" is ticked then "model estimation" should be ticked as well}\ali{fixed.}}
+  \caption{Software processing steps (adapted from~\cite{bowring2019exploring}).}
   \label{table:pipeline-steps}
@@
 
-\subsection{Fuzzy Libmath environment}
+\subsection{Within-tool variability simulation with Fuzzy Libmath}
 
-To introduce numerical noise in the analyses, we used
+We simulated within-tool variability by introducing 
+numerical noise in the analyses using
 Fuzzy Libmath~\cite{salari2021accurate}, a version of the GNU
@@
 }
-\lstinputlisting[caption=Sample wrapper function (C code) \gk{I don't think we need this/it adds anything.
-Could go to a supplement if we want it in a paper somewhere}, label=algo:wrapper, style=customasm]{wrapper.c}
+\lstinputlisting[caption=Sample wrapper function (C code),
+  % \gk{I don't think we need this/it adds anything. Could go to a supplement
+  % if we want it in a paper somewhere} \tristan{I agree it's a bit raw, but
+  % I find it gives a more concrete idea of how the whole instrumentation
+  % works. Otherwise we might as well keep equation (1) only, but there's
+  % quite a bit of work between the equation and pipeline instrumentation. It
+  % also shows how minimally invasive the instrumentation is.}
+  label=algo:wrapper, style=customasm]{wrapper.c}
 
@@
 version 18.1.09, and SPM12 version r7771 executed with GNU/Octave version
-5.2. These software versions were identical to the study
+5.2. These software versions were identical to the ones used 
 in~\cite{bowring2019exploring} except for SPM for which we used GNU/Octave
@@
 in~\cite{bowring2019exploring} were single-threaded. This was meant to
-reduce the time overhead resulting from Fuzzy Libmath instrumentation
-\gk{Did we verify that single-threading generally didn't lead to different results than 7-threading? As in, were they all ostensibly sampled from the same random distribution?}.
-\ali{We verified that multithreading leads to different results (e.g. single-threading vs, 7-threading).
-  However, single-threaded executions couldn't reproduce the original results.}
+reduce the time overhead resulting from Fuzzy Libmath instrumentation.
+% \gk{Did we verify that single-threading generally didn't lead to different
+% results than 7-threading? As in, were they all ostensibly sampled from the
+% same random distribution?} \ali{We verified that multithreading leads to
+% different results (e.g. single-threading vs, 7-threading). However,
+% single-threaded executions couldn't reproduce the original results.}
+% \tristan{Fig S1 answers this comment and is refered later on}
 All the analyses were conducted on the CentOS 7.3 operating system. The
@@
 
-We measured numerical (within-tool; WT) variability by running the same analyses three
+We simulated within-tool (WT) variability by running the same analyses three
 times using Fuzzy Libmath with a virtual precision of $t=53$~bits for
@@
 error. The resulting samples were equally plausible estimates of
-the true \yohan{How can you assert that when you don't have ground truth?
-Would it be more safe to say "plausible estimates numerical solutions"?} numerical result at the precision used by the pipelines.
-Moreover, to evaluate numerical perturbations at different magnitudes,
+the true
+% \yohan{How can you assert that when you don't have ground truth? Would it be more safe to say "plausible estimates numerical solutions"?}
+% \tristan{This does not refer to the ground truth but to the true \emph{numerical} result}
+numerical result at the precision used by the pipelines. Moreover, to evaluate numerical perturbations at different magnitudes,
 we repeated the FSL analyses for virtual
@@
 
-We evaluated WT and BT variability for thresholded as well as unthresholded
-group-level and subject-level t-statistics maps by computing the
-absolute differences of t-statistic maps across pipelines or Fuzzy Libmath samples, respectively.
-For BT, we computed the absolute differences across a given pair of tools.
-For WT, we averaged the absolute differences between three pair combinations across the three
-Fuzzy Libmath samples of a given tool. We computed WT variability for a pair of tools (A, B) as follows:
-\begin{equation}
-  WT(A,B) = \frac{WT(A) + WT(B)}{2},
-  \label{eq:wt-pair}
-\end{equation}
-where $WT(.)$ is the numerical variability for a given tool
-and $WT(., .)$ is the numerical variability for a pair of tools.
+We evaluated BT variability for thresholded as well as unthresholded
+group-level and subject-level t-statistics maps by computing the absolute
+differences of t-statistic maps across tools. For WT, we computed the
+average absolute difference across the three Fuzzy Libmath samples.
+Moreover, we computed WT variability for a pair of tools as the sum of WT values for each tool.
+\tristan{Ali, can you update the figures to (1) show WT variability for each tool instead of for tool pairs (2) remove difference between BT and WT (3) show the XY plot as a joint histogram?}
 
@@
 pair of tools (A, B) if for tool A or tool B it contained activated voxels only for some Fuzzy Libmath
-samples. \yohan{A voxel is activated if its intensity is $>$ 0? If so, the given definition focuses on regions with low intensities while regions with high intensities may be more unstable}
+samples.
+% \yohan{A voxel is activated if its intensity is $>$ 0? If so, the given definition focuses on regions with low intensities while regions with high intensities may be more unstable}
+% \tristan{A voxel is activated if its value (t-stat) is above a threshold}
 
@@
 
-Table~\ref{table:pipeline-stats} presents summary statistics for the
+Table~\ref{table:pipeline-stats} presents summary statistics for variability in the
 group-level t-statistics. For each tool pair (A, B), BT variability was
@@
 these global differences were confirmed regionally.
-In addition, BT and WT variability appeared moderately correlated across voxels (Pearson's R
-in [0.60, 0.62], p=0.0, Figure~\ref{fig:unthresh-maps}-\textbf{D}). 
+In addition, BT and WT variability appeared moderately correlated across voxels (Pearson's r
+in [0.60, 0.62], p$< 10^{-5}$, Figure~\ref{fig:unthresh-maps}-\textbf{D}). 
 % Interestingly, the correlation appears driven by a set of voxels exhibiting comparable BT and
@@
 BT in some regions. As for the group maps, BT and WT
-variability appeared moderately correlated for all subjects (R in [0.484,
-    0.633], p=0.0 for all subjects, see Supplemental Material~\ref{sec:supp-subjects}).
+variability appeared moderately correlated for all subjects (r in [0.484,
+    0.633], p$< 10^{-5}$ for all subjects, see Supplemental Material~\ref{sec:supp-subjects}).
 
@@
 between BT and WT in unthresholded group maps.
-FSL produced group maps with $\mu=0.483$ and $\sigma=0.410$ at this virtual precision,
-which indicates a bit lower variability compared to BT variability
+FSL produced unthresholded group maps with $\mu=0.483$ and $\sigma=0.410$ at this virtual precision,
+which was still lower than BT variability
 (Wilcoxon signed-rank test and t-test p~\textless~$10^{-5}$ for all tests).
-Figure~\ref{fig:gnp-mni} shows BT
-and WT in the corresponding group maps obtained at t=17 bits, showing that
-BT and WT reach comparable magnitudes in some regions. BT and WT remained
-moderately correlated at this precision.
+However, Figure~\ref{fig:gnp-mni} shows that
+BT and WT reached comparable magnitudes in some regions. BT and WT remained
+moderately correlated at this precision \tristan{add r score}.
 
@@
 that introduced variability at the edges of
-active regions for both BT and WT. Except in these regions, BT remained consistently larger
-than WT \gk{what regions? This kind of came from nowhere}. Moreover, to measure correlation between BT and WT, we
+active regions for both BT and WT. Except at the edges, BT remained consistently larger
+than WT.
+%\gk{what regions? This kind of came from nowhere}\tristan{I meant "at the edges", fixed.}
+ Moreover, to measure correlation between BT and WT, we
 measured WT instability and BT instability in each region of the

### 2021-12-22 paper/paper.tex
Update all figures

@@
   % \fbox{\begin{minipage}{\dimexpr \textwidth-2\fboxsep-2\fboxrule}
-  \begin{subfigure}[ht]{0.9\textwidth}
+  \begin{subfigure}[ht]{.9\textwidth}
     \centering
-    \includegraphics[width=0.9\textwidth]{figures/abs/gl-unthresh.png}
+    \includegraphics[width=.9\textwidth]{figures/gl-unthresh.png}
+    % \includegraphics[width=0.9\textwidth]{figures/abs/gl-unthresh.png}
     %\caption{Standard deviation of thresholded t-statistics map on template surface}
   \end{subfigure}
-  \caption{Unthresholded group-level variability computed between tools
-    (\textbf{A}), within tools at machine error (\textbf{B}), difference
-    between them (\textbf{C}), and voxel-wise comparison (\textbf{D}). }
+  \caption{Unthresholded group-level differences computed between tools
+    (\textbf{A}), within tools at machine error (\textbf{B}),
+    and voxel-wise comparison of absolute differences (\textbf{C}). }
   \label{fig:unthresh-maps}
@@
   \centering
-  \includegraphics[width=0.9\textwidth]{figures/abs/sbj05-abs.png}
+  \includegraphics[width=0.9\textwidth]{figures/plots/Bland-Altman/unthresh-sbj05.png}
+  % \includegraphics[width=0.9\textwidth]{figures/abs/sbj05-abs.png}
   %\caption{Standard deviation of thresholded t-statistics map on template surface}
   \caption{For subject with highest WT variability,
-    unthresholded subject-level variability computed between tools (\textbf{A}), within tools at machine error (\textbf{B}),
-    and difference between them (\textbf{C}).
+    unthresholded subject-level variability computed between tools (\textbf{A}), and within tools at machine error (\textbf{B}).}
+    % and difference between them (\textbf{C}).
 \camille{If the changes are visible, I think that it would be useful to add the correspondant unthresholded maps for the reader to see where the differences are and what they "look like"}}
@@
 
-%%%%%%%%%% plot different precisions%%%%%%%%
+%%%%%%%%%% plot at precision t=17%%%%%%%%
 \begin{figure*}[ht]
@@
     \centering
-    \includegraphics[width=.75\textwidth]{figures/bg_global_precision.pdf}
+    \includegraphics[width=.55\textwidth]{figures/p17-diffs.pdf}
+    % \includegraphics[width=.75\textwidth]{figures/bg_global_precision.pdf}
     %\caption{Standard deviation of thresholded t-statistics map on template surface}
@@
   %\caption{Standard deviation of thresholded t-statistics map on template surface}
-  \caption{Unthresholded group t-statistics absolute differences computed between tools (\textbf{A}),
-    within tools at the virtual precision of t=17~bits (\textbf{B}), difference between them (\textbf{C}), and
-    voxel-wise comparison (\textbf{D}).}
+  \caption{Unthresholded group t-statistics differences computed
+  within tools at the virtual precision of t=17~bits (\textbf{A}),
+  and voxel-wise comparison of absolute differences between BT and WT (\textbf{B}).}
   \label{fig:gnp-mni}
@@
     \centering
-    \includegraphics[width=\textwidth]{figures/abs/gl-thresh.png}
+    \includegraphics[width=\textwidth]{figures/plots/wt-tstats/act_deact-framed.png}
+    % \includegraphics[width=\textwidth]{figures/abs/gl-thresh.png}
     %\caption{Standard deviation of thresholded t-statistics map on template surface}
@@
   \centering
-  \caption{Thresholded group-level variability computed between tools (\textbf{A}),
-    within tools at machine error (\textbf{B}), difference between them (\textbf{C}), and
-    confusion matrices of activation instability
+  \caption{Thresholded group-level t-statistics within tools at machine error for FSL (\textbf{A}),
+    SPM (\textbf{B}), AFNI (\textbf{C}), and confusion matrices of activation instability
     in BT and WT among the 360 regions of the HCP-MMP1.0 parcellation (\textbf{D}).}
@@
   \centering
-  \includegraphics[width=.6\textwidth]{figures/sbj-abs-corr-unthresh-plot.png}
+  \includegraphics[width=.55\textwidth]{figures/sbj-abs-corr-unthresh-plot.png}
+  % \includegraphics[width=.6\textwidth]{figures/sbj-abs-corr-unthresh-plot.png}
   \caption{Comparison between BT and WT variabilty for 16 subjects.}
@@
 
+\section{Maps of t-statistics for subject with highest WT variability}
+\label{sec:supp-worst-subject}
+
+Figure~\ref{fig:unthresh-worst-sbj} plots the maps of t-statistics for each run of tools in WT.
+
+\begin{figure*}[ht]
+  \centering
+  \includegraphics[width=\textwidth]{figures/plots/wt-tstats/5-framed.png}
+  % \includegraphics[width=.6\textwidth]{figures/sbj-abs-corr-unthresh-plot.png}
+  \caption{For subject with highest WT variability, unthresholded subject-level t-statistics
+  within tools at machine error for FSL (\textbf{A}),
+  SPM (\textbf{B}), AFNI (\textbf{C}).}
+  \label{fig:unthresh-worst-sbj}
+\end{figure*}
+
 \end{document}

### 2021-12-22 paper/paper.tex
Fix figures and refs

@@
 study in~\cite{bowring2019exploring} compared the results produced by the
-three main fMRI analysis toolboxes, namely SPM, AFNI and FSL, using similar
+three main fMRI analysis toolboxes, namely SPM~\cite{penny2011statistical},
+AFNI~\cite{cox1996afni} and FSL~\cite{jenkinson2012fsl}, using similar
 pipelines. It reported limited similarity between the activation clusters
@@
 in~\cite{Li2021.12.01.470790} also showed a low similarity between the
-results produced by five different tools (ABCD, CCS, CPAC, DPARSF, fMRIPrep
-\tristan{Ali, could you add refs to these tools from Li et al 2021? Also
-for SPM, FSL and AFNI in the previous sentence}) and identified the main
+results produced by five different tools (ABCD~\cite{feczko2021adolescent},
+CCS~\cite{xu2015connectome}, CPAC~\cite{craddock2013towards}, DPARSF~\cite{yan2010dparsf},
+fMRIPrep~\cite{esteban2019fmriprep}) and identified the main
 factors contributing to these differences. The magnitude of the highlighted
@@
 Moreover, we computed WT variability for a pair of tools as the sum of WT values for each tool.
-\tristan{Ali, can you update the figures to (1) show WT variability for each tool instead of for tool pairs (2) remove difference between BT and WT (3) show the XY plot as a joint histogram?}
 
@@
 % for all tests, t-test p=0 \gk{similar comment re p value} for all tests) \camille{p is smaller than a very small p but not zero?}.
-Figure~\ref{fig:unthresh-maps}-\textbf{A},\textbf{B},\textbf{C} shows that
+Figure~\ref{fig:unthresh-maps}-\textbf{A},\textbf{B} shows that
 these global differences were confirmed regionally.
 In addition, BT and WT variability appeared moderately correlated across voxels (Pearson's r
-in [0.60, 0.62], p$< 10^{-5}$, Figure~\ref{fig:unthresh-maps}-\textbf{D}). 
+in [0.60, 0.62], p$< 10^{-5}$, Figure~\ref{fig:unthresh-maps}-\textbf{C}). 
 % Interestingly, the correlation appears driven by a set of voxels exhibiting comparable BT and
@@
 variability appeared moderately correlated for all subjects (r in [0.484,
-    0.633], p$< 10^{-5}$ for all subjects, see Supplemental Material~\ref{sec:supp-subjects}).
+    0.633], p$< 10^{-5}$ for all subjects, see Supplemental Material~\ref{sec:supp-subjects} and~\ref{sec:supp-worst-subject}).
 
@@
 BT and WT reached comparable magnitudes in some regions. BT and WT remained
-moderately correlated at this precision \tristan{add r score}.
+moderately correlated at this precision (Pearson's r
+in [0.53, 0.61], p$< 10^{-5}$, Figure~\ref{fig:gnp-mni}-\textbf{B}).
 
@@
     \centering
-    \includegraphics[width=\textwidth]{figures/plots/wt-tstats/act_deact-framed.png}
+    \includegraphics[width=\textwidth]{figures/act_deact-marked.pdf}
+    % \includegraphics[width=\textwidth]{figures/plots/wt-tstats/act_deact-framed.png}
     % \includegraphics[width=\textwidth]{figures/abs/gl-thresh.png}
@@
 
-Figure~\ref{fig:unthresh-worst-sbj} plots the maps of t-statistics for each run of tools in WT.
+Figure~\ref{fig:unthresh-worst-sbj} plots the maps of t-statistics
+for each run of tools for the subject with highest WT variability.
 
@@
   \centering
-  \includegraphics[width=\textwidth]{figures/plots/wt-tstats/5-framed.png}
+  \includegraphics[width=\textwidth]{figures/5-marked.pdf}
+  % \includegraphics[width=\textwidth]{figures/plots/wt-tstats/5-framed.png}
   % \includegraphics[width=.6\textwidth]{figures/sbj-abs-corr-unthresh-plot.png}

### 2021-12-23 paper/paper.tex
A pass on intro-methods-results

@@
 
-\title{Comparing variability across and within fMRI analysis software packages}
+\title{Comparing software variability across and within fMRI analysis packages}
 
@@
 \begin{abstract}
-  Variability has been broadly observed in functional MRI analyses as a result of differences
+  Variability has been broadly observed in functional MRI analyses as a result
+  of software differences
   both between and within analytic tools. However, the
-  relationship between within-tool and between-tool variabilities for a
+  relationship between within-tool and between-tool software variabilities for a
   given analysis is unclear. We extended a previous comparison of fMRI
   analysis software libraries (namely FSL, AFNI, and SPM) and related the
-  observed differences to within-tool variability
+  observed differences to within-tool software variability
   simulated using Monte-Carlo arithmetic. In group analyses, we found that
-  between-tool variability was consistently larger than within-tool
-  variability simulated from numerical perturbations of the magnitude of
-  machine error. In subject analyses, the same within-tool variability
-  approached between-tool variability in some regions for some subjects.
-  Interestingly, within- and between-tool variabilities appeared moderately
+  between-tool software variability was consistently larger than within-tool software
+  variability simulated at the magnitude of
+  machine error. In subject analyses, the same within-tool software variability
+  approached between-tool software variability in some regions for some subjects.
+  Interestingly, within- and between-tool software variabilities appeared moderately
   correlated ($r > 0.5$) in all tested conditions. Finally, we found that
-  within-tool variability had effects of similar magnitude to between-tool
-  variability when simulated from numerical perturbations introduced at the virtual
+  within-tool software variability had effects of similar magnitude to between-tool
+  software variability when simulated from numerical perturbations introduced at the virtual
   precision of 17 bits --- a high but plausible level of uncertainty for
-  scientific computing libraries \gk{What does this mean? They only have
-  precision of 17, or they use data that is 17 bits?}.
-   \yohan{agree with Greg, it is unclear and a strong assertion!}
-   \camille{It is not entirely clear to me how this relates to the previous
-   conclusions that BT variability was consistently larger than machine
-   error} \tristan{Better?} Our findings motivate the
-  continued investigation of within-tool variability in neuroimaging, and position it
-  as a possible proxy for studying arbitrary between-tool variations.
+  scientific computing libraries. \tristan{Revise from latest results}
+  %  \gk{What does this mean? They only have
+  % precision of 17, or they use data that is 17 bits?}
+  %  \yohan{agree with Greg, it is unclear and a strong assertion!}
+  %  \camille{It is not entirely clear to me how this relates to the previous
+  %  conclusions that BT variability was consistently larger than machine
+  %  error} \tristan{Better?}
+   Our findings motivate the
+  continued investigation of within-tool software variability in neuroimaging, and position it
+  as a possible proxy for studying some of the arbitrary between-tool software variations.
 \end{abstract}
@@
 complex software stacks that leverage low-level libraries provided by the
-operating system (e.g.,mathematical functions), general scientific
+operating system (e.g., mathematical functions), general scientific
 computing methods (e.g., optimization toolboxes), and specific fMRI
@@
 different fMRI analysis tools implementing similar analytical approaches
-(between-tool variability) can reach worrying magnitudes. For instance, the
+(between-tool software variability), or different versions of the same tool,
+ can reach worrying magnitudes. For instance, the
 study in~\cite{bowring2019exploring} compared the results produced by the
@@
 factors contributing to these differences. The magnitude of the highlighted
-differences suggest that between-tool variability may be playing a critical
+differences suggest that between-tool software variability may be playing a critical
 role in the reproducibility of fMRI analysis overall.
@@
 The variability resulting from differences in lower-level software
-libraries (within-tool variability) has also been quantified in fMRI. The
+libraries (within-tool software variability) has also been quantified in fMRI. The
 study in~\cite{Glatard2015} mentions a low similarity between the activation clusters
@@
 
-The relationship between within- and between-tool variability are poorly
+The relationship between within- and between-tool software variability are poorly
 understood, but could play an important role in the construction of
@@
 
-This paper investigates the relationship between numerical stability -- used
-as a proxy for within-tool variability -- and between-tool variability
+This paper investigates the relationship between numerical stability --- used
+as a proxy for within-tool software variability --- and between-tool software variability
 through two main questions:
 \begin{enumerate}
-\item what is the relative magnitude of within-tool variability and between-tool variability, and
-\item is there an association between within- and between-tool variability.
+\item what is the relative magnitude of within- and between-tool software variability, and
+\item is there an association between within- and between-tool software variability.
 \end{enumerate}
@@
 
-\subsection{Within-tool variability simulation with Fuzzy Libmath}
+\subsection{Within-tool software variability simulation with Fuzzy Libmath}
 
-We simulated within-tool variability by introducing 
+We simulated within-tool software variability by introducing 
 numerical noise in the analyses using
@@
 
-We measured between-tool (BT) variability by running the pipelines
+We measured between-tool software variability (BT) by running the pipelines
 described in~\cite{bowring2019exploring} with FSL version 5.0.10, AFNI
@@
 we encapsulated the
-above-mentioned software packages in Docker container images based on CentOS 7.3.
-Although, we converted Docker containers to Singularity images to enable running on cluster nodes.
+above-mentioned software packages in Docker container images based on CentOS 7.3
+which we converted to Singularity images to enable running on cluster nodes.
 
-We simulated within-tool (WT) variability by running the same analyses three
+We simulated within-tool software variability (WT) by running the same analyses three
 times using Fuzzy Libmath with a virtual precision of $t=53$~bits for
@@
 values were chosen such that the numerical perturbation simulates machine
-error. The resulting samples were equally plausible estimates of
+error. The resulting samples are equally plausible estimates of
 the true
@@
 
-We evaluated BT variability for thresholded as well as unthresholded
+We evaluated BT for thresholded as well as unthresholded
 group-level and subject-level t-statistics maps by computing the absolute
@@
 
-\subsection{Validation of Replication}
+\subsection{Validation of replication}
 
@@
 
-\subsection{In the group analysis, BT variability was larger than WT variability}
+\subsection{In the group analysis, BT was larger than WT}
 
 Table~\ref{table:pipeline-stats} presents summary statistics for variability in the
-group-level t-statistics. For each tool pair (A, B), BT variability was
+group-level t-statistics. For each tool pair (A, B), BT was
 significantly larger than machine error in tool A or B, in both thresholded
@@
 % for all tests, t-test p=0 \gk{similar comment re p value} for all tests) \camille{p is smaller than a very small p but not zero?}.
-Figure~\ref{fig:unthresh-maps}-\textbf{A},\textbf{B} shows that
-these global differences were confirmed regionally.
-In addition, BT and WT variability appeared moderately correlated across voxels (Pearson's r
+These global differences were confirmed by Bland-Altman plots showing a
+clear dominance of BT over WT
+(Figure~\ref{fig:unthresh-maps}-\textbf{A},\textbf{B}).
+In addition, BT and WT appeared moderately correlated across voxels (Pearson's r
 in [0.60, 0.62], p$< 10^{-5}$, Figure~\ref{fig:unthresh-maps}-\textbf{C}). 
@@
   \end{subfigure}
-  \caption{Unthresholded group-level differences computed between tools
-    (\textbf{A}), within tools at machine error (\textbf{B}),
-    and voxel-wise comparison of absolute differences (\textbf{C}). }
+  \caption{\textbf{A} and \textbf{B}: Bland-Altman plots comparing group-level differences computed between tools
+    (\textbf{A}) and within tools at machine error (\textbf{B}). \textbf{C}: voxel-wise comparison of absolute differences.}
   \label{fig:unthresh-maps}
@@
 
-\subsection{In subject analyses, WT variability approached BT variability in some regions}
+\subsection{In subject analyses, WT approached BT for some subjects}
 
@@
 subject-level unthresholded t-statistics maps. As for group-level maps,
-BT variability was consistently larger than WT variability (Wilcoxon
+BT was consistently larger than WT (Wilcoxon
 signed-rank test and t-test p~\textless~$10^{-5}$ for all tests).
-However, Figure~\ref{fig:unthresh-maps-sbj} shows that for
-some subjects, WT variability approached %and even surpassed
-BT in some regions. As for the group maps, BT and WT
-variability appeared moderately correlated for all subjects (r in [0.484,
+However, for
+some subjects and for AFNI, WT approached 
+BT in some regions (Figure~\ref{fig:unthresh-maps-sbj}). As for the group maps, BT and WT
+ appeared moderately correlated for all subjects (r in [0.484,
     0.633], p$< 10^{-5}$ for all subjects, see Supplemental Material~\ref{sec:supp-subjects} and~\ref{sec:supp-worst-subject}).
@@
 
-\subsection{At precision t=17~bits, WT variability approached BT variability in the group analysis}
-
-While the previous results were obtained for machine error, we also
+\subsection{At precision t=17~bits and in FSL, WT approached BT in the group analysis}
+\tristan{We cannot say that from the data that we have.}
+\tristan{Discuss subject-level BA plots.}
+While the previous results were obtained by simulating WT with numerical perturbations 
+at the magnitude of machine error, we also
 evaluated numerical variability across different virtual precisions for FSL
@@
 
-Figure~\ref{fig:thresh-maps}-\textbf{A},\textbf{B},\textbf{C} compares BT
-and WT for thresholded group maps. Thresholding is an unstable operation
+We also compared BT and WT in thresholded maps since these maps are commonly used instead of unthresholded ones 
+to conclude on the activation of specific regions. Thresholding is an unstable operation
 that introduced variability at the edges of
 active regions for both BT and WT. Except at the edges, BT remained consistently larger
-than WT.
-%\gk{what regions? This kind of came from nowhere}\tristan{I meant "at the edges", fixed.}
- Moreover, to measure correlation between BT and WT, we
-measured WT instability and BT instability in each region of the
-HCP-MMP1.0 parcellation \gk{again, it feels kind of abrupt and I'm not really sure what this section is getting at or how it differs from the previous based on these descriptions.
-Perhaps introduce that thresholding is common, but since it makes things discontinuous we study it through aparcellation, etc...}.
-The confusion matrices in
-Figure~\ref{fig:thresh-maps}-\textbf{D} report these instabilities for
-the 360 tested regions. The average ratio of unstable regions was 26\%
-for BT and 9.2\% for WT, which confirmed that BT variability was larger than machine error.
+than WT (Figure~\ref{fig:thresh-maps}-\textbf{A},\textbf{B},\textbf{C} and Table~\ref{table:pipeline-stats}).
+
+Moreover, to compare the effects of BT and WT on the detection of activated
+regions, we measured WT instability and BT instability in each region of
+the HCP-MMP1.0 parcellation. The confusion matrices in
+Figure~\ref{fig:thresh-maps}-\textbf{D} report these instabilities for the
+360 tested regions. The average ratio of unstable regions was 26\% for BT
+and 9.2\% for WT, which confirmed that BT was larger than
+WT even in thresholded maps.
+
+Finally, we measured agreement between BT and WT from the confusion
+matrices since interpreting the correlation of thresholded maps is
+difficult due to the discontinuity introduced by thresholding. 
 The average Cohen's kappa score\footnote{$\kappa \leq 0$ denotes chance agreement, $-1 \leq \kappa \leq 1$}
-between WT instability and BT instability was $\kappa=0.17$, indicating
-a moderate agreement between WT instability and BT instability.
+computed from the confusion matrices between WT instability and BT instability was $\kappa=0.17$, indicating
+a moderate agreement between WT instability and BT instability consistent with the correlations observed in 
+unthresholded maps.
+
+% \gk{again, it feels kind of abrupt and I'm not really sure what this
+% section is getting at or how it differs from the previous based on these
+% descriptions. Perhaps introduce that thresholding is common, but since it
+% makes things discontinuous we study it through aparcellation, etc...}.
+% \tristan{better now?}
 
@@
 introduced by our replication due to the use of GNU/Octave vs MATLAB for SPM
-and multithreading for AFNI \camille{Would this be feasible to check?}. However, the group maps remained very similar
+and multithreading for AFNI.
+ \camille{Would this be feasible to check?}
+ \tristan{Ali, could you answer that?}
+  However, the group maps remained very similar
 overall, which led us to conclude that our results correctly reproduced
@@
   \caption{Differences between reproduced and original results obtained in~\cite{bowring2019exploring}
-    of unthresholded group-level t-statistics for SPM (left) and AFNI (right). \camille{it looks like the highest areas of difference in AFNI are due to difference un masking maybe worth noting?}}
+    of unthresholded group-level t-statistics for SPM (left) and AFNI
+    (right). The highest areas of difference in AFNI seem to be due to
+    differences in brain masks.}
   \label{fig:replication-diff}

### 2021-12-23 paper/paper.tex

@@
   \centering
-  \caption{Thresholded group-level t-statistics within tools at machine error for FSL (\textbf{A}),
-    SPM (\textbf{B}), AFNI (\textbf{C}), and confusion matrices of activation instability
-    in BT and WT among the 360 regions of the HCP-MMP1.0 parcellation (\textbf{D}).}
+  \caption{\textbf{A},\textbf{B},\textbf{C}: Thresholded group-level
+    t-statistics within tools at machine error for FSL (\textbf{A}), SPM
+    (\textbf{B}) and AFNI (\textbf{C}). Arrows point to activation clusters
+    impacted by both within- and between-tool variability \tristan{Ali: is it really the case? Can you make the arrows point more accurately to 
+    the clusters? Some of them are quite far from the cluster they point to.}.
+     \textbf{D}: Confusion matrices of
+    activation instability in BT and WT among the 360 regions of the
+    HCP-MMP1.0 parcellation.}
   \label{fig:thresh-maps}
@@
 
-In fMRI group analyses, within-tool variability remains an order of magnitude smaller
-than between-tool variability. Group analyses benefit from regularization
-of numerical noise, which is expected to increase as sample size
-increases. Therefore, for fMRI studies with large sample sizes, within-tool variability may be negligible with respect to between-tool variability. The
-recommendation made in~\cite{botvinik2020variability} to rely on
-``multiverse'' analyses where multiple analysis tools are compared is
-therefore likely to successfully correct for machine error in group
-studies. Nevertheless, within-tool variability remains substantial in group analyses
-that are based on a single tool, as is commonly the case in current fMRI
-studies. In particular, in our study, the inherent instability of
-thresholding was triggered by machine error in 9\% of 360 brain regions,
-which indicates that machine error might have impacted neuroscientific
-conclusions related to these regions.
+In fMRI group analyses, within-tool software variability remains an order
+of magnitude smaller than between-tool variability. This is likely
+explained by the fact that group analyses benefit from regularization of
+numerical noise, which is expected to increase with sample size. This finding is
+consistent with observations made in~\cite{kiar2020numerical} from
+diffusion MRI data where connectome graph statistics were found to be
+substantially unstable at the subject level while group distributions
+remained consistent. Therefore,
+for fMRI studies with large sample sizes, within-tool software variability
+may be neglected with respect to between-tool variability. In particular,
+multiverse analyses aggregating the outcome of multiple analysis tools are
+likely to successfully correct for machine error in such group studies.  Nevertheless, within-tool software variability remains
+substantial in group analyses that are based on a single tool, as is
+commonly the case in current fMRI studies. In particular, in our study, the
+inherent instability of thresholding was triggered by within-tool software
+variability in 9\% of 360 brain regions, which indicates that it might have
+impacted neuroscientific conclusions related to these regions.
 
-In subject-level analyses, within-tool and between-tool variabilty can become
+In subject-level analyses, within- and between-tool software variabilties can become
 of comparable magnitude for some subjects in some regions. This observation
 is particularly relevant to the development of fMRI-based biomarkers aiming
-at individualized phenotype predictions. Machine error is expected to play
+at individualized phenotype predictions. Machine error may play
 a non-negligible role in such analyses, even when predictions combine
-results produced by multiple tools. The observed regularization effect of
-group analyses toward numerical noise is consistent with observations made
-in~\cite{kiar2020numerical} from diffusion MRI where connectome graph
-statistics were found to be substantially unstable at the subject-level
-while group distributions remained consistent. \gk{this last sentence feels like it belongs in the previous paragraph, right before mentioning thresholding}
+results produced by multiple tools.
 
-For both group- and subject-level analyses, between-tool variability and within-tool
-variability were found to be moderately correlated. Even though between-tool
-variability and within-tool  variability are different in nature, this result
-suggests that in some cases both sources of variability may have a common
-cause that might be related to the conditioning of the BOLD signal
-estimation problem \gk{I wouldn't necessarily say it's because of "BOLD signal estimation problem", just "algorithms used to process"} in a specific dataset. Therefore, in some regions,
-instabilities of similar magnitude may be triggered by small numerical
-perturbations, model variations, or implementation differences. For
-instance, the analysis of datasets with high motion may be unstable both
-across tools and numerically.
+For both group- and subject-level analyses, between- and within-tool
+software variabilities were found to be moderately correlated. Even though
+both types of variability are different in nature, this result suggests
+that in some cases they may have a common cause that might be related to
+the conditioning of fMRI analysis in a specific dataset. Along these lines,
+in some regions, instabilities of similar magnitude may be triggered by
+small numerical perturbations, model variations, or implementation
+differences. For instance, high-motion datasets may trigger between- and
+within-tool instabilities similarly.
 
@@
 between-tool variability. This speculation might be of practical value to
-address software variability at large given that numerical stability refers
-to a consistent mathematical framework whereas between-tool variability
-remains more empirical. Likewise, useful quality control metrics may be derived from numerical and between-tool stability.
+address software variability at large given that within-tool variability
+measurements can be automated while between-tool ones require substantial
+human intervention. Indeed, within-tool variability can be simulated
+irrespective of the particular analysis done, and with important sample
+sizes, while between-tool variability requires re-implementing analysis in
+different toolboxes and is limited to a few samples. Automated multi-verse
+frameworks could be developed to assess and possibly reduce within-tool
+variability as a pre-requisite for between-tool variability.
 
+\tristan{This paragraph needs to be revised as we can't conclude that from the data we have}
 The finding that numerical variability approached tool variability at the
 virtual precision of t=17~bits is interesting too. Indeed, while machine
-error generally introduces differences in the order of \yohanmod{1} \yohan{1/2} ulp
-\yohan{Are we talking about one arithmetic operation? Or an entire execution?}
---- or
-t=53~bits for double-precision values and t=24~bits for single-precision
-ones \yohan{I would not mix virtual precision and error measurement}--- common scientific software dependencies introduce larger errors
+error generally introduces perturbations on the least-significant bit of the mantissa,
+ones 
+% \yohan{I would not mix virtual precision and error measurement}\tristan{better?}
+common scientific software dependencies introduce larger perturbations
 .
 For instance, SciPy's 2D spline interpolation was recently found to be
-precise up to t=10~bits~\cite{pytracer} \tristan{Fix ref}
-\yohan{Same as previous, I would talk about significance or relative error, or remove the t= and not talk about virtual precision} and might therefore introduce
+precise up to 10~bits~\cite{pytracer}
+and its replacement by a more precise implementation might therefore introduce
 numerical perturbations leading to errors in the range of between-tool
-variability \tristan{Yohan, please check that sentence}. Such high errors
+variability. Such high errors
 might be triggered by updates in operating systems, Python, MATLAB, and
 other software dependencies.
-\yohan{The sentence is ok, just a picky point: the SciPy result you cited describes the precision of the spline implementation while at the end you're talking differences due to updates.
-  The nuance is that two versions may have a similar precision and so when you udpate your
-  system, the difference between both versions is low but each version can be ill-conditionned.}
+% \yohan{The sentence is ok, just a picky point: the SciPy result you cited
+% describes the precision of the spline implementation while at the end
+% you're talking differences due to updates. The nuance is that two versions
+% may have a similar precision and so when you udpate your system, the
+% difference between both versions is low but each version can be
+% ill-conditionned.}
+% \tristan{better?}
 
 Our results are limited by the type of numerical noise introduced in the
-analyses. We only perturbed the outputs of elementary
+analyses. Indeed, we only perturbed the outputs of elementary
 mathematical functions while numerical noise could creep in any
-floating-point operation. Our estimation of numerical variability should
-therefore be considered a lower bound. Our estimation of tool varability is
-also likely to be underestimated, having tested only 3 analytical pipelines
+floating-point operation. Therefore, our estimation of within-tool software variability should
+be considered a lower bound. Likewise, our estimation of tool varability is
+likely to be underestimated, having tested only 3 analytical pipelines
 among the thousands available~\cite{carp2012plurality}.
@@
 In conclusion, our results motivate further numerical stability
-investigations in fMRI analysis pipelines. Pipeline-level analyses could be
-conducted to identify specific components that contribute to numerical
-variability, and if possible correct them accordingly. When instability is
-inherent to the analysis, sampling results distributions through numerical
-perturbations might improve stability, as explored in~\cite{kiar2021data}.
-Finally, pipeline-specific statistical corrections might be envisaged to
-account for numerical variability.
+investigations in fMRI analyses. Pipeline-level analyses could be conducted
+to identify specific components that contribute to numerical variability,
+and if possible correct them accordingly. Besides, in cases where
+instability is inherent to the analysis, sampling results distributions
+through numerical perturbations might improve stability, as explored
+in~\cite{kiar2021data}. Finally, pipeline-specific statistical corrections
+might be envisaged to account for within-tool software variability.
 

### 2021-12-23 paper/paper.tex
Add subject-level BA plots at p17

@@
     impacted by both within- and between-tool variability \tristan{Ali: is it really the case? Can you make the arrows point more accurately to 
-    the clusters? Some of them are quite far from the cluster they point to.}.
+    the clusters? Some of them are quite far from the cluster they point to.}\ali{Yes, it's the case.}.
      \textbf{D}: Confusion matrices of
@@
  \tristan{Ali, could you answer that?}
+ \ali{Yes, all the output data are available in Beluga cluster in `~/projects/def-glatard/between-tool-samples/results/ds000001/'.
+ These replicated results can find in particular in `tools/AFNI/replicated' for AFNI and `tools/SPM/Octave' for SPM results.}
   However, the group maps remained very similar

### 2021-12-31 paper/paper.tex
Update results at t=17bits

--- 2021-12-23 paper/paper.tex
@@
   Interestingly, within- and between-tool software variabilities appeared moderately
-  correlated ($r > 0.5$) in all tested conditions. Finally, we found that
-  within-tool software variability had effects of similar magnitude to between-tool
-  software variability when simulated from numerical perturbations introduced at the virtual
-  precision of 17 bits --- a high but plausible level of uncertainty for
-  scientific computing libraries. \tristan{Revise from latest results}
-  %  \gk{What does this mean? They only have
+  correlated ($r > 0.5$) in all tested conditions. 
+  % Finally, we found that
+  % within-tool software variability had effects of similar magnitude to between-tool
+  % software variability when simulated from numerical perturbations introduced at the virtual
+  % precision of 17 bits --- a high but plausible level of uncertainty for
+  % scientific computing libraries. \tristan{Revise from latest results}
+  % %  \gk{What does this mean? They only have
   % precision of 17, or they use data that is 17 bits?}
@@
 
-\subsection{At precision t=17~bits and in FSL, WT approached BT in the group analysis}
-\tristan{We cannot say that from the data that we have.}
-\tristan{Discuss subject-level BA plots.}
-While the previous results were obtained by simulating WT with numerical perturbations 
-at the magnitude of machine error, we also
-evaluated numerical variability across different virtual precisions for FSL
-and found that the virtual precision of t=17 bits minimized the RMSE
-between BT and WT in unthresholded group maps.
-FSL produced unthresholded group maps with $\mu=0.483$ and $\sigma=0.410$ at this virtual precision,
-which was still lower than BT variability
-(Wilcoxon signed-rank test and t-test p~\textless~$10^{-5}$ for all tests).
-However, Figure~\ref{fig:gnp-mni} shows that
-BT and WT reached comparable magnitudes in some regions. BT and WT remained
-moderately correlated at this precision (Pearson's r
-in [0.53, 0.61], p$< 10^{-5}$, Figure~\ref{fig:gnp-mni}-\textbf{B}).
+% \subsection{At precision t=17~bits and in FSL, WT approached BT in the group analysis}
+% \tristan{We cannot say that from the data that we have.}
+% \tristan{Discuss subject-level BA plots.}
+% While the previous results were obtained by simulating WT with numerical perturbations 
+% at the magnitude of machine error, we also
+% evaluated numerical variability across different virtual precisions for FSL
+% and found that the virtual precision of t=17 bits minimized the RMSE
+% between BT and WT in unthresholded group maps.
+% FSL produced unthresholded group maps with $\mu=0.483$ and $\sigma=0.410$ at this virtual precision,
+% which was still lower than BT variability
+% (Wilcoxon signed-rank test and t-test p~\textless~$10^{-5}$ for all tests).
+% However, Figure~\ref{fig:gnp-mni} shows that
+% BT and WT reached comparable magnitudes in some regions. BT and WT remained
+% moderately correlated at this precision (Pearson's r
+% in [0.53, 0.61], p$< 10^{-5}$, Figure~\ref{fig:gnp-mni}-\textbf{B}).
 
-%%%%%%%%%% plot at precision t=17%%%%%%%%
-\begin{figure*}[ht]
-  \begin{subfigure}[ht]{\textwidth}
-    \centering
-    \includegraphics[width=.55\textwidth]{figures/p17-diffs.pdf}
-    % \includegraphics[width=.75\textwidth]{figures/bg_global_precision.pdf}
-    %\caption{Standard deviation of thresholded t-statistics map on template surface}
-  \end{subfigure}
-  %\caption{Standard deviation of thresholded t-statistics map on template surface}
-  \caption{Unthresholded group t-statistics differences computed
-  within tools at the virtual precision of t=17~bits (\textbf{A}),
-  and voxel-wise comparison of absolute differences between BT and WT (\textbf{B}).}
-  \label{fig:gnp-mni}
-\end{figure*}
+% %%%%%%%%%% plot at precision t=17%%%%%%%%
+% \begin{figure*}[ht]
+%   \begin{subfigure}[ht]{\textwidth}
+%     \centering
+%     \includegraphics[width=.55\textwidth]{figures/p17-diffs.pdf}
+%     % \includegraphics[width=.75\textwidth]{figures/bg_global_precision.pdf}
+%     %\caption{Standard deviation of thresholded t-statistics map on template surface}
+%   \end{subfigure}
+%   %\caption{Standard deviation of thresholded t-statistics map on template surface}
+%   \caption{Unthresholded group t-statistics differences computed
+%   within tools at the virtual precision of t=17~bits (\textbf{A}),
+%   and voxel-wise comparison of absolute differences between BT and WT (\textbf{B}).}
+%   \label{fig:gnp-mni}
+% \end{figure*}
 
@@
     (\textbf{B}) and AFNI (\textbf{C}). Arrows point to activation clusters
-    impacted by both within- and between-tool variability \tristan{Ali: is it really the case? Can you make the arrows point more accurately to 
-    the clusters? Some of them are quite far from the cluster they point to.}.
+    impacted by both within- and between-tool variability.
      \textbf{D}: Confusion matrices of
@@
 
-\tristan{This paragraph needs to be revised as we can't conclude that from the data we have}
-The finding that numerical variability approached tool variability at the
-virtual precision of t=17~bits is interesting too. Indeed, while machine
-error generally introduces perturbations on the least-significant bit of the mantissa,
-ones 
-% \yohan{I would not mix virtual precision and error measurement}\tristan{better?}
-common scientific software dependencies introduce larger perturbations
-.
-For instance, SciPy's 2D spline interpolation was recently found to be
-precise up to 10~bits~\cite{pytracer}
-and its replacement by a more precise implementation might therefore introduce
-numerical perturbations leading to errors in the range of between-tool
-variability. Such high errors
-might be triggered by updates in operating systems, Python, MATLAB, and
-other software dependencies.
+% \tristan{This paragraph needs to be revised as we can't conclude that from the data we have}
+% The finding that numerical variability approached tool variability at the
+% virtual precision of t=17~bits is interesting too. Indeed, while machine
+% error generally introduces perturbations on the least-significant bit of the mantissa,
+% ones 
+% % \yohan{I would not mix virtual precision and error measurement}\tristan{better?}
+% common scientific software dependencies introduce larger perturbations
+% .
+% For instance, SciPy's 2D spline interpolation was recently found to be
+% precise up to 10~bits~\cite{pytracer}
+% and its replacement by a more precise implementation might therefore introduce
+% numerical perturbations leading to errors in the range of between-tool
+% variability. Such high errors
+% might be triggered by updates in operating systems, Python, MATLAB, and
+% other software dependencies.
 % \yohan{The sentence is ok, just a picky point: the SciPy result you cited
@@
 and multithreading for AFNI.
- \camille{Would this be feasible to check?}
- \tristan{Ali, could you answer that?}
+%  \camille{Would this be feasible to check?}
+%  \tristan{Ali, could you answer that?}
+%  \ali{Yes, all the output data are available in Beluga cluster in `~/projects/def-glatard/between-tool-samples/results/ds000001/'.
+%  These replicated results can find in particular in `tools/AFNI/replicated' for AFNI and `tools/SPM/Octave' for SPM results.}
   However, the group maps remained very similar

### 2022-01-02 paper/paper.tex
Minor edits on method

@@
 error. The resulting samples are equally plausible estimates of
-the true
+the true numerical result at the precision used by the pipelines. 
 % \yohan{How can you assert that when you don't have ground truth? Would it be more safe to say "plausible estimates numerical solutions"?}
 % \tristan{This does not refer to the ground truth but to the true \emph{numerical} result}
-numerical result at the precision used by the pipelines. Moreover, to evaluate numerical perturbations at different magnitudes,
-we repeated the FSL analyses for virtual
-precisions ranging from $t=1$~bit to $t=24$~bits for single-precision
-and double-precision values. We also repeated the FSL analyses for virtual
-precisions ranging from $t=24$~bits to $t=53$~bits for double-precision values, having
-set the virtual precision to $t=24$~bits for single-precision values.
+% Moreover, to evaluate numerical perturbations at different magnitudes,
+% we repeated the FSL analyses for virtual
+% precisions ranging from $t=1$~bit to $t=24$~bits for single-precision
+% and double-precision values. We also repeated the FSL analyses for virtual
+% precisions ranging from $t=24$~bits to $t=53$~bits for double-precision values, having
+% set the virtual precision to $t=24$~bits for single-precision values.
 

### 2022-01-03 paper/paper.tex
camille comments on abstract

@@
   relationship between within-tool and between-tool software variabilities for a
-  given analysis is unclear. We extended a previous comparison of fMRI
+  given analysis is unclear \camille{early on in the abstract, I think that it would be worth specifying what within-tool covers in the context of this paper: i.e.
+a focus on variations that can be induced by low level libraries as seen for example when using different operating systems -- which is different from within-tool variability that could be induced when using different parameters (e.g smoothing 6mm or 8mm) or algorithms (e.g.
+FLIRT versus FNIRT in FSL) within the same software } .
+We extended a previous comparison of fMRI
   analysis software libraries (namely FSL, AFNI, and SPM) and related the
@@
   Interestingly, within- and between-tool software variabilities appeared moderately
-  correlated ($r > 0.5$) in all tested conditions. 
+  correlated ($r > 0.5$) \camille{maybe add ``across voxels or across xxx'' so that we understand what the correlation is measuring a bit better?} in all tested conditions . 
   % Finally, we found that
@@
   as a possible proxy for studying some of the arbitrary between-tool software variations.
+\camille{should we add ``for subject-level analyses'' here?
+Since we said earlier that within-tool was consistently smaller than between tool variability for group studies.}
 \end{abstract}

### 2022-01-03 paper/paper.tex
comment methods

@@
 differences suggest that between-tool software variability may be playing a critical
-role in the reproducibility of fMRI analysis overall.
+role in the reproducibility of \camille{add ``an'' ?} fMRI analysis overall.
 
@@
   \end{tabular}
-  \caption{Software processing steps (adapted from~\cite{bowring2019exploring}).}
+  \caption{Software processing steps (adapted from~\cite{bowring2019exploring}).
+\camille{``Analysis Voxel Size'' is not a processing step per se but more a parametrization of inter-subject registration so maybe it should be removed?
+// Not sure if this is worth noting but in FSL ``Inter-subject registration'' happens after the first-level analysis}}
   \label{table:pipeline-steps}
@@
 5.2. These software versions were identical to the ones used 
-in~\cite{bowring2019exploring} except for SPM for which we used GNU/Octave
+in~\cite{bowring2019exploring} except for SPM for which we \camille{add ``had to use a more recent version in order to be able to use it with''} GNU/Octave
 instead of MATLAB to enable mathematical function instrumentation using
@@
 values were chosen such that the numerical perturbation simulates machine
-error. The resulting samples are equally plausible estimates of
+error \camille{Given the target journal where any readers (like myself :) ) won't be expert in nuemrical perturbations, I think that you could add a bit more details here on why those numbers are good to simulate machine errors}.
+The resulting samples are equally plausible estimates of
 the true numerical result at the precision used by the pipelines. 

### 2022-01-03 paper/paper.tex
fix typo?

@@
 We evaluated BT for thresholded as well as unthresholded
-group-level and subject-level t-statistics maps by computing the absolute
+group-level and subject-level t-statistic maps by computing the absolute
 differences of t-statistic maps across tools. For WT, we computed the

### 2022-01-03 paper/paper.tex
camille comment results start

@@
 \section{Results}
-The scripts, Docker images, and data to reproduce the results are available
+The scripts, Docker images, and data \camille{maybe a detail but consider using something like ``and derived datasets that were used to reproduce the results'' ?
+As it is now, it made me wonder whether the whole nifti dataset was in the github repo} to reproduce the results are available
 in our GitHub repository at

### 2022-01-03 paper/paper.tex
camille on repro

@@
 in~\cite{bowring2019exploring}. For FSL, we found
-identical results. For SPM, the files (as measured by checksums) were different but differences
+identical results \camille{It would be nice here to be more specific about how you checked that they were identical -- using a sha sum ?
+And maybe even insisting by using ``striclty identifical''.
+I think for different people in the community ``identical'' can mean different things}.
+For SPM, the files (as measured by checksums) were different but differences
 were visually unnoticeable. For AFNI, the activation maps were similar
 overall, however, differences were noticeable visually.
-The observed differences remained small (see Supplemental Material~\ref{sec:supp-repro}), and might be due to the use of
+The observed differences remained small \camille{maybe also add ``and the overall patterns of activations were preseved''} (see Supplemental Material~\ref{sec:supp-repro}), and might be due to the use of
 GNU/Octave vs MATLAB in SPM, and of multithreading in AFNI. We performed visual quality control of the AFNI

### 2022-01-03 paper/paper.tex
camille more comments

@@
 Table~\ref{table:pipeline-stats} presents summary statistics for variability in the
-group-level t-statistics. For each tool pair (A, B), BT was
+group-level t-statistics.
+For each tool pair (A, B), BT \camille{could we restate which metric we are looking at here -- as it stands BT means between-tool but coul represent many different metrics?
+It is ``absolute difference BT''?} was
 significantly larger than machine error in tool A or B, in both thresholded
@@
   \caption{Voxel-wise mean and standard deviation of BT and WT variability
-    in t-statistics maps.}
+    in t-statistics maps \camille{here and in other places -- is it t-statistic maps? (i.e. without s after statistic)}.}
   \label{table:pipeline-stats}
@@
 Table~\ref{table:pipeline-stats} also presents summary statistics for
-subject-level unthresholded t-statistics maps. As for group-level maps,
+subject-level unthresholded t-statistics maps \camille{Is this for one subject in particular ? If so how was it chosen ?Or a mean over all subjects?}. As for group-level maps,
 BT was consistently larger than WT (Wilcoxon
@@
 Figure~\ref{fig:unthresh-correlation-allsbj} plots the relationship between
-WT and BT for each subject, showing a consistent moderate correlation between BT and
+WT and BT for each subject \camille{across voxels?}, showing a consistent moderate correlation between BT and
 WT.
@@
   within tools at machine error for FSL (\textbf{A}),
-  SPM (\textbf{B}), AFNI (\textbf{C}).}
+  SPM (\textbf{B}), AFNI (\textbf{C}).
+\camille{If I understand correctly the point you are making here is that WT(AFNI) at some voxels is close to AFNI-SPM ?
+What is surprising here for me is that WT(AFNI) seems to be quite different from WT(SPM) and WT(FSL).
+Do we have a good idea why?
+Maybe looking at the maps would help?} }
   \label{fig:unthresh-worst-sbj}

### 2022-01-10 paper/paper.tex
Apply changes from Camille

--- 2022-01-03 paper/paper.tex
@@
 differences suggest that between-tool software variability may be playing a critical
-role in the reproducibility of \camille{add ``an'' ?} fMRI analysis overall.
+role in the reproducibility of an fMRI analysis overall.
 
@@
     {}                     & {Inter-subject Registration}       & \checkmark & \checkmark & \checkmark \\
-    {}                     & {Analysis Voxel Size}              & \checkmark & \checkmark & \checkmark \\
+    % {}                     & {Analysis Voxel Size}              & \checkmark & \checkmark & \checkmark \\
     {}                     & {Smoothing}                        & \checkmark & \checkmark & \checkmark \\
@@
   \end{tabular}
-  \caption{Software processing steps (adapted from~\cite{bowring2019exploring}).
-\camille{``Analysis Voxel Size'' is not a processing step per se but more a parametrization of inter-subject registration so maybe it should be removed?
+  \caption{Software processing steps (adapted from~\cite{bowring2019exploring}).}
+  % \camille{``Analysis Voxel Size'' is not a processing step per se but more a parametrization of inter-subject registration so maybe it should be removed?
 // Not sure if this is worth noting but in FSL ``Inter-subject registration'' happens after the first-level analysis}}
@@
 5.2. These software versions were identical to the ones used 
-in~\cite{bowring2019exploring} except for SPM for which we \camille{add ``had to use a more recent version in order to be able to use it with''} GNU/Octave
+in~\cite{bowring2019exploring} except for SPM for which we had to use
+a more recent version in order to be able to use it with GNU/Octave
 instead of MATLAB to enable mathematical function instrumentation using
@@
 values were chosen such that the numerical perturbation simulates machine
-error \camille{Given the target journal where any readers (like myself :) ) won't be expert in nuemrical perturbations, I think that you could add a bit more details here on why those numbers are good to simulate machine errors}.
+error, an upper bound on the relative error between a real number and its floating-point representation.
+% \camille{Given the target journal where any readers (like myself :) ) won't be expert in nuemrical perturbations, I think that you could add a bit more details here on why those numbers are good to simulate machine errors}.
 The resulting samples are equally plausible estimates of
@@
 \section{Results}
-The scripts, Docker images, and data \camille{maybe a detail but consider using something like ``and derived datasets that were used to reproduce the results'' ?
-As it is now, it made me wonder whether the whole nifti dataset was in the github repo} to reproduce the results are available
-in our GitHub repository at
-\url{https://github.com/big-data-lab-team/fuzzy-neurotools}.
+The scripts, Docker images, and derived datasets that were used to reproduce the results are available
+in our GitHub repository at \url{https://github.com/big-data-lab-team/fuzzy-neurotools}.
 
@@
 in~\cite{bowring2019exploring}. For FSL, we found
-identical results. For SPM, the files (as measured by checksums) were different but differences
-were visually unnoticeable. For AFNI, the activation maps were similar
-overall, however, differences were noticeable visually.
+identical results with the same checksums. For SPM, the files (as measured by checksums) were different but differences
+were visually unnoticeable. For AFNI, the overall patterns of activations were preseved
+, however, differences were noticeable visually.
 The observed differences remained small (see Supplemental Material~\ref{sec:supp-repro}), and might be due to the use of
@@
 Table~\ref{table:pipeline-stats} presents summary statistics for variability in the
-group-level t-statistics. For each tool pair (A, B), BT was
+group-level t-statistic. For each tool pair (A, B), absolute difference BT was
 significantly larger than machine error in tool A or B, in both thresholded
@@
   \caption{Voxel-wise mean and standard deviation of BT and WT variability
-    in t-statistics maps.}
+    in t-statistic maps.}
   \label{table:pipeline-stats}
@@
 Table~\ref{table:pipeline-stats} also presents summary statistics for
-subject-level unthresholded t-statistics maps. As for group-level maps,
-BT was consistently larger than WT (Wilcoxon
+subject-level unthresholded t-statistic maps computed by the mean across all subjects.
+As for group-level maps, BT was consistently larger than WT (Wilcoxon
 signed-rank test and t-test p~\textless~$10^{-5}$ for all tests).
@@
   \caption{\textbf{A},\textbf{B},\textbf{C}: Thresholded group-level
-    t-statistics within tools at machine error for FSL (\textbf{A}), SPM
+    t-statistic within tools at machine error for FSL (\textbf{A}), SPM
     (\textbf{B}) and AFNI (\textbf{C}). Arrows point to activation clusters
@@
   \caption{Differences between reproduced and original results obtained in~\cite{bowring2019exploring}
-    of unthresholded group-level t-statistics for SPM (left) and AFNI
+    of unthresholded group-level t-statistic for SPM (left) and AFNI
     (right). The highest areas of difference in AFNI seem to be due to
@@
 Figure~\ref{fig:unthresh-correlation-allsbj} plots the relationship between
-WT and BT for each subject, showing a consistent moderate correlation between BT and
+WT and BT for each subject across voxels, showing a consistent moderate correlation between BT and
 WT.
@@
 
-\section{Maps of t-statistics for subject with highest WT variability}
+\section{Maps of t-statistic for subject with highest WT variability}
 \label{sec:supp-worst-subject}
 
-Figure~\ref{fig:unthresh-worst-sbj} plots the maps of t-statistics
+Figure~\ref{fig:unthresh-worst-sbj} plots the maps of t-statistic
 for each run of tools for the subject with highest WT variability.
@@
   % \includegraphics[width=.6\textwidth]{figures/sbj-abs-corr-unthresh-plot.png}
-  \caption{For subject with highest WT variability, unthresholded subject-level t-statistics
+  \caption{For subject with highest WT variability, unthresholded subject-level t-statistic
   within tools at machine error for FSL (\textbf{A}),
-  SPM (\textbf{B}), AFNI (\textbf{C}).}
+  SPM (\textbf{B}), AFNI (\textbf{C}).
+\camille{If I understand correctly the point you are making here is that WT(AFNI) at some voxels is close to AFNI-SPM ?
+What is surprising here for me is that WT(AFNI) seems to be quite different from WT(SPM) and WT(FSL).
+Do we have a good idea why?
+Maybe looking at the maps would help?} }
   \label{fig:unthresh-worst-sbj}

### 2022-01-18 paper/paper.tex
Fixes in abstract

@@
 \begin{abstract}
-  Variability has been broadly observed in functional MRI analyses as a result
-  of software differences
-  both between and within analytic tools. However, the
-  relationship between within-tool and between-tool software variabilities for a
-  given analysis is unclear \camille{early on in the abstract, I think that it would be worth specifying what within-tool covers in the context of this paper: i.e.
-a focus on variations that can be induced by low level libraries as seen for example when using different operating systems -- which is different from within-tool variability that could be induced when using different parameters (e.g smoothing 6mm or 8mm) or algorithms (e.g.
-FLIRT versus FNIRT in FSL) within the same software } .
-We extended a previous comparison of fMRI
-  analysis software libraries (namely FSL, AFNI, and SPM) and related the
-  observed differences to within-tool software variability
-  simulated using Monte-Carlo arithmetic. In group analyses, we found that
+  Variability has been broadly observed in functional MRI analyses as a
+  result of software differences both between and within analytic tools.
+  However, the relationship between within-tool and between-tool software
+  variabilities for a given analysis is unclear. Within-tool variability has multiple origins
+  including variations in algorithms, parametrizations, and
+  low-level software dependencies. We focus on the effects resulting from software dependencies --- 
+  in particular due to operating system and other infrastructural updates --- as they often remain unnoticed 
+  while algorithms and parametrizations are more controllable.
+ We extended
+  a previous comparison of fMRI analysis software libraries (namely FSL,
+  AFNI, and SPM) and related the observed differences to within-tool
+  software variability simulated using Monte-Carlo arithmetic. In group
+  analyses, we found that
   between-tool software variability was consistently larger than within-tool software
-  variability simulated at the magnitude of
-  machine error. In subject analyses, the same within-tool software variability
+  variability. In subject analyses, within-tool software variability
   approached between-tool software variability in some regions for some subjects.
   Interestingly, within- and between-tool software variabilities appeared moderately
-  correlated ($r > 0.5$) \camille{maybe add ``across voxels or across xxx'' so that we understand what the correlation is measuring a bit better?} in all tested conditions . 
+  correlated ($r > 0.5$ across voxels) in all tested conditions. 
   % Finally, we found that
@@
   %  error} \tristan{Better?}
-   Our findings motivate the
-  continued investigation of within-tool software variability in neuroimaging, and position it
-  as a possible proxy for studying some of the arbitrary between-tool software variations.
-\camille{should we add ``for subject-level analyses'' here?
-Since we said earlier that within-tool was consistently smaller than between tool variability for group studies.}
+  Our findings motivate the continued investigation of within-tool
+  software variability in neuroimaging, and position it as a possible proxy
+  for studying some of the arbitrary between-tool software variations for
+  subject-level analyses.
 \end{abstract}
@@
 Table~\ref{table:pipeline-stats} also presents summary statistics for
-subject-level unthresholded t-statistic maps computed by the mean across all subjects.
+subject-level unthresholded t-statistic maps (mean over subjects).
 As for group-level maps, BT was consistently larger than WT (Wilcoxon

### 2022-01-18 paper/paper.tex
Update minor changes

@@
 
-\author{Ali Salari$^1$, Yohan Chatelain$^1$, Alexander Bowring$^2$, Camille Maumet$^3$, Gregory Kiar$^4$, Tristan Glatard$^1$\\
+\author{Ali Salari$^1$, Yohan Chatelain$^1$, Camille Maumet$^3$, Gregory Kiar$^4$, Tristan Glatard$^1$\\
   \vspace*{0.1cm}\\
   $^1$ Department of Computer-Science and Software Engineering, Concordia University, Montreal, Canada\\
-  $^2$ Li Ka Shing Centre for Health Information and Discovery, Nuffield Department of Population Health,\\ Big Data Institute, University of Oxford, Oxford, UK\\
+  % $^2$ Li Ka Shing Centre for Health Information and Discovery, Nuffield Department of Population Health,\\ Big Data Institute, University of Oxford, Oxford, UK\\
   $^3$ Inria, Univ Rennes, CNRS, Inserm, IRISA UMR 6074, Empenn ERL U 1228, Rennes, France\\
@@
 
+A floating-point number can be represented by a finite number of precision bits, 24 bits
+for single-precision and 53 bits for double-precision in the IEEE-754 standard.
+Thus, the machine error, the smallest difference between two floating-point numbers that the computer
+can recognize, for single-precision is $2^{-24}$ and double-precision is $2^{-53}$.
+We simulated the machine error by instrumenting the last precision bits of the floating-point numbers.
 We simulated within-tool software variability (WT) by running the same analyses three
 times using Fuzzy Libmath with a virtual precision of $t=53$~bits for
-double-precision values and $t=24$~bits for single-precision values. These
-values were chosen such that the numerical perturbation simulates machine
-error, an upper bound on the relative error between a real number and its floating-point representation.
+double-precision values and $t=24$~bits for single-precision values. 
+These values were chosen such that the numerical perturbation simulates machine error.
+%an upper bound on the relative error between a real number and its floating-point representation.
 % \camille{Given the target journal where any readers (like myself :) ) won't be expert in nuemrical perturbations, I think that you could add a bit more details here on why those numbers are good to simulate machine errors}.
@@
 in~\cite{bowring2019exploring}. For FSL, we found
-identical results with the same checksums. For SPM, the files (as measured by checksums) were different but differences
+strictly identical results with the same MD5 checksums. For SPM, the files (as measured by checksums) were different but differences
 were visually unnoticeable. For AFNI, the overall patterns of activations were preseved
@@
   within tools at machine error for FSL (\textbf{A}),
-  SPM (\textbf{B}), AFNI (\textbf{C}).
-\camille{If I understand correctly the point you are making here is that WT(AFNI) at some voxels is close to AFNI-SPM ?
+  SPM (\textbf{B}), AFNI (\textbf{C}).}% \camille{If I understand correctly the point you are making here is that WT(AFNI) at some voxels is close to AFNI-SPM ?
 What is surprising here for me is that WT(AFNI) seems to be quite different from WT(SPM) and WT(FSL).
 Do we have a good idea why?
-Maybe looking at the maps would help?} }
+Maybe looking at the maps would help?}
+  %\ali{We are pointing regions with instability in both WT and BT, and yes indeed it shows a very different WT(AFNI) compare to WT(FSL) and WT(SPM), but we did not investigate the reason of variabilities for now.} }
   \label{fig:unthresh-worst-sbj}

### 2022-01-24 paper/paper.tex
Comments in section 2.3

@@
 
-A floating-point number can be represented by a finite number of precision bits, 24 bits
-for single-precision and 53 bits for double-precision in the IEEE-754 standard.
-Thus, the machine error, the smallest difference between two floating-point numbers that the computer
-can recognize, for single-precision is $2^{-24}$ and double-precision is $2^{-53}$.
-We simulated the machine error by instrumenting the last precision bits of the floating-point numbers.
+A floating-point number is represented by a finite number of bits, 24 bits
+for single-precision and 53 bits for double-precision in the IEEE-754 standard \tristan{This is the size of the mantissas, not the size of the entire representation}.
+Machine error, the smallest difference between two floating-point numbers that can be represented,
+is $2^{-24}$ in single-precision and $2^{-53}$ in double precision \tristan{The magnitude of machine error depends on the exponent}.
+We simulated machine error by instrumenting the last precision bits \tristan{We cannot ``instrument'' a bit, we can perturb it. Code is instrumented.}
+ of the floating-point numbers.
 We simulated within-tool software variability (WT) by running the same analyses three
@@
 in~\cite{bowring2019exploring}. For FSL, we found
-strictly identical results with the same MD5 checksums. For SPM, the files (as measured by checksums) were different but differences
+strictly identical results with the same MD5 checksums. For SPM, the files (as measured by MD5 checksums) were different but differences
 were visually unnoticeable. For AFNI, the overall patterns of activations were preseved

### 2022-01-24 paper/paper.tex
Added Alex back

@@
 
-\author{Ali Salari$^1$, Yohan Chatelain$^1$, Camille Maumet$^3$, Gregory Kiar$^4$, Tristan Glatard$^1$\\
+\author{Ali Salari$^1$, Yohan Chatelain$^1$, Alexander Bowring$^2$, Camille Maumet$^3$, Gregory Kiar$^4$, Tristan Glatard$^1$\\
   \vspace*{0.1cm}\\
   $^1$ Department of Computer-Science and Software Engineering, Concordia University, Montreal, Canada\\
-  % $^2$ Li Ka Shing Centre for Health Information and Discovery, Nuffield Department of Population Health,\\ Big Data Institute, University of Oxford, Oxford, UK\\
+  $^2$ Li Ka Shing Centre for Health Information and Discovery, Nuffield Department of Population Health,\\ Big Data Institute, University of Oxford, Oxford, UK\\
   $^3$ Inria, Univ Rennes, CNRS, Inserm, IRISA UMR 6074, Empenn ERL U 1228, Rennes, France\\

### 2022-01-25 paper/paper.tex
Minor changes

@@
 
-A floating-point number is represented by a finite number of bits, 24 bits
-for single-precision and 53 bits for double-precision in the IEEE-754 standard \tristan{This is the size of the mantissas, not the size of the entire representation}.
-Machine error, the smallest difference between two floating-point numbers that can be represented,
-is $2^{-24}$ in single-precision and $2^{-53}$ in double precision \tristan{The magnitude of machine error depends on the exponent}.
-We simulated machine error by instrumenting the last precision bits \tristan{We cannot ``instrument'' a bit, we can perturb it. Code is instrumented.}
- of the floating-point numbers.
+In the IEEE-754 format, a floating-point number $N$ is represented by a sign
+bit (s), an exponent (e), and a pseudo-mantissa (m) as:
+
+\begin{equation}
+  N = (-1)^s \times 2^E \times (1.m),
+\end{equation}
+where $E=e-127$ for single precision and $E=e-1023$ for double precision.
+In single precision, the pseudo-mantissa has 23 bits and the exponent has 8 bits, whereas in 
+double precision the pseudo-mantissa has 52 bits and the exponent has 11 bits. 
+Machine error, the smallest absolute difference between two
+floating-point numbers that can be represented, is in general $2^{-24}\times e$ in
+single precision and $2^{-53}\times e$ in double precision.
+We simulated machine error by introducing the random perturbation defined
+in Equation~\ref{eq} at the virtual precision of $t=24$~bits for single-precision 
+values and $t=53$~bits for double-precision ones. The perturbed pseudo-mantissa
+is obtained by extending the size of the original pseudo-mantissa, applying the perturbation 
+at the desired virtual precision, and rounding the perturbed result
+ back to the original precision.  
 We simulated within-tool software variability (WT) by running the same analyses three

### 2022-01-25 paper/paper.tex
Fix ref

@@
 We simulated machine error by introducing the random perturbation defined
-in Equation~\ref{eq} at the virtual precision of $t=24$~bits for single-precision 
+in Equation~\ref{eq:mca_inexact} at the virtual precision of $t=24$~bits for single-precision 
 values and $t=53$~bits for double-precision ones. The perturbed pseudo-mantissa

### 2022-02-08 paper/paper.tex
Update results

@@
 is obtained by extending the size of the original pseudo-mantissa, applying the perturbation
-at the desired virtual precision, and rounding the perturbed result
- back to the original precision.  
+at the desired virtual precision, and rounding the perturbed result back to the original precision.
 We simulated within-tool software variability (WT) by running the same analyses three
@@
 double-precision values and $t=24$~bits for single-precision values. 
-These values were chosen such that the numerical perturbation simulates machine error.
-%an upper bound on the relative error between a real number and its floating-point representation.
-% \camille{Given the target journal where any readers (like myself :) ) won't be expert in nuemrical perturbations, I think that you could add a bit more details here on why those numbers are good to simulate machine errors}.
+Perturbations were applied at the same magnitude of floating-point precisions during the pipeline execution
+regardless of what it corresponds to.
+These values were chosen such that the numerical perturbation affects all the pipeline steps and simulates
+machine error originating from software, hardware, and operating system changes.
 The resulting samples are equally plausible estimates of
 the true numerical result at the precision used by the pipelines.
-% \yohan{How can you assert that when you don't have ground truth? Would it be more safe to say "plausible estimates numerical solutions"?}
-% \tristan{This does not refer to the ground truth but to the true \emph{numerical} result}
-% Moreover, to evaluate numerical perturbations at different magnitudes,
-% we repeated the FSL analyses for virtual
-% precisions ranging from $t=1$~bit to $t=24$~bits for single-precision
-% and double-precision values. We also repeated the FSL analyses for virtual
-% precisions ranging from $t=24$~bits to $t=53$~bits for double-precision values, having
-% set the virtual precision to $t=24$~bits for single-precision values.
 
 We evaluated BT for thresholded as well as unthresholded
-group-level and subject-level t-statistic maps by computing the absolute
+group-level and subject-level t-statistic maps by computing the
 differences of t-statistic maps across tools. For WT, we computed the
-average absolute difference across the three Fuzzy Libmath samples.
-Moreover, we computed WT variability for a pair of tools as the sum of WT values for each tool.
+average difference across the three Fuzzy Libmath samples.
+Moreover, we computed WT variability for a pair of tools as the sum of WT absolute differences for each tool.
+All differences were computed over the intersection of t-statistic maps between the pair of tools.
 
@@
 Table~\ref{table:pipeline-stats} presents summary statistics for variability in the
-group-level t-statistic. For each tool pair (A, B), absolute difference BT was
-significantly larger than machine error in tool A or B, in both thresholded
-and unthresholded maps (Wilcoxon signed-rank test and t-test p~\textless~$10^{-5}$ for all tests).
-% \gk{we don't have sufficient sample size to say that p is that small,
-% so we should probably say less than $10^{-5}$ or something else we could actually
-% arrive at with our number of repetitions, pipeline pairs, and samples.}
-% for all tests, t-test p=0 \gk{similar comment re p value} for all tests) \camille{p is smaller than a very small p but not zero?}.
+group-level t-statistic. 
+Mean of differences in t-statistic maps is generally centered around zero for BT and WT
+(one-sample t-test p~\textless~$10^{-4}$ for all tests).
+However, for each tool pair (A, B), BT variability was
+significantly larger than machine error in tool A or B,
+with an order of magnitude larger standard deviation of differences 
+in both thresholded and unthresholded maps, 
+(Wilcoxon signed-rank test and t-test p~\textless~$10^{-4}$ for all tests).
 These global differences were confirmed by Bland-Altman plots showing a
@@
 In addition, BT and WT appeared moderately correlated across voxels (Pearson's r
-in [0.60, 0.62], p$< 10^{-5}$, Figure~\ref{fig:unthresh-maps}-\textbf{C}). 
+in [0.68, 0.69], p$< 10^{-5}$, Figure~\ref{fig:unthresh-maps}-\textbf{C}). 
 % Interestingly, the correlation appears driven by a set of voxels exhibiting comparable BT and
@@
     \rowcolor{lightgray!50}
-    {Between Tools}   & FSL vs. SPM  & 2.564                           & 1.050                              & 0.887                             & 0.689    & 0.733 & 0.586    \\
+    {Between Tools}   & FSL vs. SPM  & -0.668                           & 1.056                              & -0.033                             & 1.121    & 0.068 & 0.937    \\
     \rowcolor{lightgray!50}
-    {(BT)}            & FSL vs. AFNI & 3.096                           & 1.232                              & 1.095                             & 0.882    & 0.878 & 0.704    \\
+    {(BT)}            & FSL vs. AFNI & -0.823                           & 1.857                              & 0.048                             & 1.439    & 0.034 & 1.146    \\
     \rowcolor{lightgray!50}
-    {}                & AFNI vs. SPM & 2.951                           & 1.344                              & 1.218                             & 0.955    & 0.983 & 0.762    \\
-    {Within Tool}     & FSL          & 0.510                           & 0.697                              & 0.127                             & 0.098    & 0.120 & 0.083    \\
-    {(WT)}            & SPM          & 0.367                           & 0.638                              & 0.085                             & 0.068    & 0.075 & 0.056    \\
-    {}                & AFNI         & 0.637                           & 0.747                              & 0.199                             & 0.202    & 0.167 & 0.197    \\
+    {}                & AFNI vs. SPM & 0.645                           & 1.428                              & 0.064                             & 1.489    & -0.065 & 1.206    \\
+    {Within Tool}     & FSL          & -0.003                           & 0.069                              & -0.010                             & 0.103    & -0.002 & 0.099    \\
+    {(WT)}            & SPM          & 0.001                           & 0.097                              & 0.000                             & 0.076    & 0.000 & 0.064    \\
+    {}                & AFNI         & 0.002                           & 0.228                              & -0.001                             & 0.178    & 0.000 & 0.178    \\
     \bottomrule
   \end{tabular}
-  \caption{Voxel-wise mean and standard deviation of BT and WT variability
-    in t-statistic maps.}
+  \caption{Voxel-wise mean and standard deviation of BT and WT variability in t-statistic maps.}
   \label{table:pipeline-stats}
@@
 As for group-level maps, BT was consistently larger than WT (Wilcoxon
-signed-rank test and t-test p~\textless~$10^{-5}$ for all tests).
-However, for
-some subjects and for AFNI, WT approached 
+signed-rank test and t-test p~\textless~$10^{-4}$ for 12 subjects).
+However, for some subjects and for AFNI, WT approached 
 BT in some regions (Figure~\ref{fig:unthresh-maps-sbj}). As for the group maps, BT and WT
- appeared moderately correlated for all subjects (r in [0.484,
-    0.633], p$< 10^{-5}$ for all subjects, see Supplemental Material~\ref{sec:supp-subjects} and~\ref{sec:supp-worst-subject}).
+ appeared moderately correlated for all subjects (r in [0.52,
+    0.69], p$< 10^{-5}$ for all subjects, see Supplemental Material~\ref{sec:supp-subjects} and~\ref{sec:supp-worst-subject}).
 
@@
   \centering
-  \includegraphics[width=0.9\textwidth]{figures/plots/Bland-Altman/unthresh-sbj05.png}
+  \includegraphics[width=0.9\textwidth]{figures/plots/Bland-Altman/unthresh-sbj10.png}
   % \includegraphics[width=0.9\textwidth]{figures/abs/sbj05-abs.png}
@@
 of comparable magnitude for some subjects in some regions. This observation
+can be used as a proxy for data quality, and
 is particularly relevant to the development of fMRI-based biomarkers aiming
@@
   \centering
-  \includegraphics[width=\textwidth]{figures/5-marked.pdf}
+  \includegraphics[width=\textwidth]{figures/10-marked.pdf}
   % \includegraphics[width=\textwidth]{figures/plots/wt-tstats/5-framed.png}

### 2022-02-21 paper/paper.tex
Update results

@@
 at the desired virtual precision, and rounding the perturbed result back to the original precision.
-We simulated within-tool software variability (WT) by running the same analyses three
+We simulated within-tool software variability (WT) by running the same analyses 10
 times using Fuzzy Libmath with a virtual precision of $t=53$~bits for
@@
 differences of t-statistic maps across tools. For WT, we computed the
-average difference across the three Fuzzy Libmath samples.
-Moreover, we computed WT variability for a pair of tools as the sum of WT absolute differences for each tool.
-All differences were computed over the intersection of t-statistic maps between the pair of tools.
+average difference across the 10 Fuzzy Libmath samples.
+% Moreover, we computed WT variability for a pair of tools as the sum of WT absolute differences for each tool.
+% All differences were computed over the intersection of t-statistic maps between the pair of tools.
 
@@
 group-level t-statistic.
-Mean of differences in t-statistic maps is generally centered around zero for BT and WT
-(one-sample t-test p~\textless~$10^{-4}$ for all tests).
+Mean of differences in t-statistic maps is centered around zero for BT and WT
+in both thresholded and unthresholded maps (t-test p~\textless~$10^{-5}$).
 However, for each tool pair (A, B), BT variability was
-significantly larger than machine error in tool A or B,
+significantly larger than WT in tool A or B,
 with an order of magnitude larger standard deviation of differences 
 in both thresholded and unthresholded maps, 
-(Wilcoxon signed-rank test and t-test p~\textless~$10^{-4}$ for all tests).
+(Wilcoxon signed-rank test and t-test p~\textless~$10^{-5}$).
 These global differences were confirmed by Bland-Altman plots showing a
@@
 (Figure~\ref{fig:unthresh-maps}-\textbf{A},\textbf{B}).
-In addition, BT and WT appeared moderately correlated across voxels (Pearson's r
-in [0.68, 0.69], p$< 10^{-5}$, Figure~\ref{fig:unthresh-maps}-\textbf{C}). 
+In addition, results show that BT and WT were not correlated across voxels
+(Pearson's r $\approx 0$, p$< 10^{-5}$, Figure~\ref{fig:unthresh-maps-corr}).
 % Interestingly, the correlation appears driven by a set of voxels exhibiting comparable BT and
@@
     \rowcolor{lightgray!50}
-    {Between Tools}   & FSL vs. SPM  & -0.668                           & 1.056                              & -0.033                             & 1.121    & 0.068 & 0.937    \\
+    {Between Tools}   & FSL vs. SPM  & -0.845                           & 2.639                              & -0.052                             & 1.122    & -0.059 & 0.911    \\
     \rowcolor{lightgray!50}
-    {(BT)}            & FSL vs. AFNI & -0.823                           & 1.857                              & 0.048                             & 1.439    & 0.034 & 1.146    \\
+    {(BT)}            & FSL vs. AFNI & -0.871                           & 3.216                              & 0.048                             & 1.405    & -0.027 & 1.097    \\
     \rowcolor{lightgray!50}
-    {}                & AFNI vs. SPM & 0.645                           & 1.428                              & 0.064                             & 1.489    & -0.065 & 1.206    \\
-    {Within Tool}     & FSL          & -0.003                           & 0.069                              & -0.010                             & 0.103    & -0.002 & 0.099    \\
-    {(WT)}            & SPM          & 0.001                           & 0.097                              & 0.000                             & 0.076    & 0.000 & 0.064    \\
-    {}                & AFNI         & 0.002                           & 0.228                              & -0.001                             & 0.178    & 0.000 & 0.178    \\
+    {}                & AFNI vs. SPM & 0.426                           & 3.214                              & -0.113                             & 1.543    & -0.104 & 1.208    \\
+    {Within Tool}     & FSL          & -0.009                           & 0.266                              & 0.012                             & 0.046    & 0.004 & 0.042    \\
+    {(WT)}            & SPM          & 0.000                           & 0.199                              & 0.000                             & 0.035    & 0.000 & 0.029    \\
+    {}                & AFNI         & -0.004                           & 0.273                              & 0.000                             & 0.084    & 0.000 & 0.075    \\
     \bottomrule
@@
   \centering
-  % \fbox{\begin{minipage}{\dimexpr \textwidth-2\fboxsep-2\fboxrule}
-  \begin{subfigure}[ht]{.9\textwidth}
+    \includegraphics[width=.9\textwidth]{figures/plots/Bland-Altman/with-masking/unthresh-gl-frame.png}
+  \caption{\textbf{A} and \textbf{B}: Bland-Altman plots comparing group-level differences computed between tools
+    (\textbf{A}) and within tools at machine error (\textbf{B}).}% \textbf{C}: voxel-wise comparison of absolute differences.}
+  \label{fig:unthresh-maps}
+\end{figure*}
+
+\begin{figure*}[ht]
   \centering
-    \includegraphics[width=.9\textwidth]{figures/gl-unthresh.png}
-    % \includegraphics[width=0.9\textwidth]{figures/abs/gl-unthresh.png}
-    %\caption{Standard deviation of thresholded t-statistics map on template surface}
-  \end{subfigure}
-  \caption{\textbf{A} and \textbf{B}: Bland-Altman plots comparing group-level differences computed between tools
-    (\textbf{A}) and within tools at machine error (\textbf{B}). \textbf{C}: voxel-wise comparison of absolute differences.}
-  \label{fig:unthresh-maps}
-  % \end{minipage}}
+  \includegraphics[width=.9\textwidth]{figures/rel-corr-unthresh-with-masking.png}
+  \caption{Voxel-wise comparison of group-level differences in BT and WT.}
+  \label{fig:unthresh-maps-corr}
 \end{figure*}
@@
 As for group-level maps, BT was consistently larger than WT (Wilcoxon
-signed-rank test and t-test p~\textless~$10^{-4}$ for 12 subjects).
+signed-rank test and t-test p~\textless~$10^{-4}$ for 14 subjects).
 However, for some subjects and for AFNI, WT approached 
-BT in some regions (Figure~\ref{fig:unthresh-maps-sbj}). As for the group maps, BT and WT
- appeared moderately correlated for all subjects (r in [0.52,
-    0.69], p$< 10^{-5}$ for all subjects, see Supplemental Material~\ref{sec:supp-subjects} and~\ref{sec:supp-worst-subject}).
+BT in some regions (Figure~\ref{fig:unthresh-maps-sbj},
+and see Supplemental Material~\ref{sec:supp-worst-subject}).
+% As for the group maps, BT and WT
+% showed no correlation across voxels for all subjects
+% (r$\approx 0$, p$< 10^{-5}$ for all subjects, see Supplemental Material~\ref{sec:supp-subjects} and~\ref{sec:supp-worst-subject}).
 
@@
   \centering
-  \includegraphics[width=0.9\textwidth]{figures/plots/Bland-Altman/unthresh-sbj10.png}
-  % \includegraphics[width=0.9\textwidth]{figures/abs/sbj05-abs.png}
-  %\caption{Standard deviation of thresholded t-statistics map on template surface}
+  \includegraphics[width=0.9\textwidth]{figures/plots/Bland-Altman/with-masking/unthresh-sbj04.png}
   \caption{For subject with highest WT variability,
@@
 Figure~\ref{fig:thresh-maps}-\textbf{D} report these instabilities for the
-360 tested regions. The average ratio of unstable regions was 26\% for BT
-and 9.2\% for WT, which confirmed that BT was larger than
+360 tested regions. The average ratio of unstable regions was 26.2\% for BT
+and 20.6\% for WT, which confirmed that BT was larger than
 WT even in thresholded maps.
@@
 The average Cohen's kappa score\footnote{$\kappa \leq 0$ denotes chance agreement, $-1 \leq \kappa \leq 1$}
-computed from the confusion matrices between WT instability and BT instability was $\kappa=0.17$, indicating
+computed from the confusion matrices between WT instability and BT instability was $\kappa=0.2$, indicating
 a moderate agreement between WT instability and BT instability consistent with the correlations observed in 
@@
     \includegraphics[width=\textwidth]{figures/act_deact-marked.pdf}
-    % \includegraphics[width=\textwidth]{figures/plots/wt-tstats/act_deact-framed.png}
-    % \includegraphics[width=\textwidth]{figures/abs/gl-thresh.png}
-    %\caption{Standard deviation of thresholded t-statistics map on template surface}
   \end{subfigure}
@@
     activation instability in BT and WT among the 360 regions of the
-    HCP-MMP1.0 parcellation.}
+    HCP-MMP1.0 parcellation.(with masking)}
   \label{fig:thresh-maps}
+\end{figure*}
+
+\subsection{Brain masking instability was triggered by WT and BT}
+
+While the previous results were obtained over the union of t-statistic maps between the pair of tools,
+we also computed results in the intersection of maps and found that the brain masking instability was
+triggered by both BT and WT.
+BT and WT variabilities were consistently smaller than ones with the fixed masks
+in both thresholded and unthresholded maps (see Table~\ref{table:pipeline-stats-fixed-mask}).
+The effect of masking was clearly appeared in BT by Bland-Altman plots showing that blob lines
+were removed after fixing masks (Figure~\ref{fig:unthresh-maps-mask}-\textbf{A},\textbf{B}).
+
+%%%%%%%%%% Summary of statstics %%%%%%%%
+\setlength{\tabcolsep}{5pt}
+\begin{table}[h]
+  \centering
+  \begin{tabular}{cccccc|cc}
+    \toprule
+                      &              & \multicolumn{4}{c|}{Group map}  & \multicolumn{2}{c}{Subject maps}                                                                     \\
+    \multirow{2}{*}{} & {}           & \multicolumn{2}{c}{Thresholded} & \multicolumn{2}{c|}{Unthresholded} & \multicolumn{2}{c}{Unthresholded}                               \\
+    % \cmidrule{3-8}
+    {}                & {}           & $\mu$                           & $\sigma$                           & $\mu$                             & $\sigma$ & $\mu$ & $\sigma$ \\
+    \midrule
+    \rowcolor{lightgray!50}
+    {Between Tools}   & FSL vs. SPM  & -0.668                           & 1.056                              & -0.033                             & 1.121    & -0.068 & 0.937    \\
+    \rowcolor{lightgray!50}
+    {(BT)}            & FSL vs. AFNI & -0.823                           & 1.857                              & 0.048                             & 1.439    & 0.034 & 1.146    \\
+    \rowcolor{lightgray!50}
+    {}                & AFNI vs. SPM & 0.645                           & 1.428                              & 0.064                             & 1.489    & -0.065 & 1.206    \\
+    {Within Tool}     & FSL          & -0.025                           & 0.197                              & 0.012                             & 0.043    & 0.004 & 0.041    \\
+    {(WT)}            & SPM          & -0.002                           & 0.154                              & 0.000                             & 0.033    & 0.000 & 0.028    \\
+    {}                & AFNI         & -0.008                           & 0.208                              & 0.000                             & 0.078    & 0.000 & 0.072    \\
+    \bottomrule
+  \end{tabular}
+  \caption{Voxel-wise mean and standard deviation of BT and WT variability in intersected t-statistic maps.}
+  \label{table:pipeline-stats-fixed-mask}
+\end{table}
+
+\begin{figure*}[ht]
+  \centering
+    \includegraphics[width=.9\textwidth]{figures/plots/Bland-Altman/fixed-mask/unthresh-gl-frame.png}
+  \caption{\textbf{A} and \textbf{B}: Bland-Altman plots comparing group-level differences computed over the intersections 
+  between tools (\textbf{A}) and within tools at machine error (\textbf{B}).}% \textbf{C}: voxel-wise comparison of absolute differences.}
+  \label{fig:unthresh-maps-mask}
 \end{figure*}
@@
 inherent instability of thresholding was triggered by within-tool software
-variability in 9\% of 360 brain regions, which indicates that it might have
+variability in 20\% of 360 brain regions, which indicates that it might have
 impacted neuroscientific conclusions related to these regions.
 
-In subject-level analyses, within- and between-tool software variabilties can become
+In subject-level analyses, within- and between-tool software variabilities can become
 of comparable magnitude for some subjects in some regions. This observation
-can be used as a proxy for data quality, and
+% can be used as a proxy for data quality, and
 is particularly relevant to the development of fMRI-based biomarkers aiming
@@
 
-\section{BT and WT correlations for all subjects}
-\label{sec:supp-subjects}
+% \section{BT and WT correlations for all subjects}
+% \label{sec:supp-subjects}
 
-Figure~\ref{fig:unthresh-correlation-allsbj} plots the relationship between
-WT and BT for each subject across voxels, showing a consistent moderate correlation between BT and
-WT.
+% Figure~\ref{fig:unthresh-correlation-allsbj} plots the relationship between
+% WT and BT for each subject across voxels, showing a consistent moderate correlation between BT and
+% WT.
 
-\begin{figure*}[ht]
-  \centering
-  \includegraphics[width=.55\textwidth]{figures/sbj-abs-corr-unthresh-plot.png}
-  % \includegraphics[width=.6\textwidth]{figures/sbj-abs-corr-unthresh-plot.png}
-  \caption{Comparison between BT and WT variabilty for 16 subjects.}
-  \label{fig:unthresh-correlation-allsbj}
-\end{figure*}
+% \begin{figure*}[ht]
+%   \centering
+%   \includegraphics[width=.55\textwidth]{figures/sbj-abs-corr-unthresh-plot.png}
+%   % \includegraphics[width=.6\textwidth]{figures/sbj-abs-corr-unthresh-plot.png}
+%   \caption{Comparison between BT and WT variabilty for 16 subjects.}
+%   \label{fig:unthresh-correlation-allsbj}
+% \end{figure*}
 
@@
   \centering
-  \includegraphics[width=\textwidth]{figures/10-marked.pdf}
+  \includegraphics[width=\textwidth]{figures/4-marked.pdf}
   % \includegraphics[width=\textwidth]{figures/plots/wt-tstats/5-framed.png}

### 2022-02-21 computs.py

@@ -15,131 +15,64 @@ import scipy.stats as stat
 from scipy.signal import resample
 import os
 import pandas as pd
+import itertools
 
 
-############# Global nearest precision
+############# COMPUTE DIFF.
 
-def calculate_rmse(im1, im2):
-    """Computing the RMSE between two images."""
-    img1 = np.nan_to_num(im1.get_fdata())
-    img2 = np.nan_to_num(im2.get_fdata())
-
-    if img1.shape != img2.shape:
-        im1 = resample_from_to(im1, im2, order=0)
-        img1 = np.nan_to_num(im1.get_fdata())
-
-    bg_ = np.where((np.isnan(im1.get_fdata())) & (np.isnan(im2.get_fdata())), False, True)
-    img1 = img1[bg_]
-    img2 = img2[bg_]
-
-    return np.sqrt(np.mean((img1-img2)**2, dtype=np.float64), dtype=np.float64)
-
-
-def global_nearest_precision():
-    rmse_ = {}
-    all_rmse = {}
-    global_average_rmse = []
-    for p in range(11, 54, 2):
-        rmse_[p] = 0
-        all_rmse[str(p)] = []
-        for s in ['fsl-afni', 'fsl-spm', 'afni-spm']:
-            bt_ = nib.load('./data/abs/{}-unthresh-abs.nii.gz'.format(s))
-            # bt_ = bt_.get_fdata()
-            wt_fsl = nib.load('./data/abs/FL-FSL/p{}_fsl_unthresh_abs.nii.gz'.format(p))
-            # wt_fsl = wt_fsl.get_fdata()
-            rmse_value = calculate_rmse(bt_, wt_fsl)
-            all_rmse[str(p)].append(rmse_value)
-            rmse_[p] += rmse_value
-
-    all_rmse = np.transpose(np.array(list(all_rmse.values())))
-    return min(rmse_, key=rmse_.get), all_rmse
-
-
-def plot_rmse_nearest(all_rmse):
-    import matplotlib.cm as cm
-    fig = plt.figure(figsize=(15,10))
-    colors = ['red', 'green', 'blue']
-    tool_pair = ['fsl-afni', 'fsl-spm', 'afni-spm']
-    for s, rmse_ in enumerate(all_rmse):
-        #type_ = ""
-        #if s+1 in i2T1w: type_ += "*"
-        #if s+1 in i2T2w: type_ += "$\dag$"
[…]

### 2022-02-21 paper/paper.tex
Update resutls with selecting max of WT samples

@@
 group-level t-statistic.
-Mean of differences in t-statistic maps is centered around zero for BT and WT
+Mean of differences in t-statistic maps was centered around zero for BT and WT
 in both thresholded and unthresholded maps (t-test p~\textless~$10^{-5}$).
@@
     \rowcolor{lightgray!50}
-    {(BT)}            & FSL vs. AFNI & -0.871                           & 3.216                              & 0.048                             & 1.405    & -0.027 & 1.097    \\
+    {(BT)}            & FSL vs. AFNI & -0.871                           & 3.216                              & 0.048                             & 1.405    & 0.027 & 1.097    \\
     \rowcolor{lightgray!50}
     {}                & AFNI vs. SPM & 0.426                           & 3.214                              & -0.113                             & 1.543    & -0.104 & 1.208    \\
-    {Within Tool}     & FSL          & -0.009                           & 0.266                              & 0.012                             & 0.046    & 0.004 & 0.042    \\
-    {(WT)}            & SPM          & 0.000                           & 0.199                              & 0.000                             & 0.035    & 0.000 & 0.029    \\
-    {}                & AFNI         & -0.004                           & 0.273                              & 0.000                             & 0.084    & 0.000 & 0.075    \\
+    {Within Tool}     & FSL          & -0.397                           & 1.218                              & -0.020                             & 0.189    & 0.017 & 0.207    \\
+    {(WT)}            & SPM          & 0.142                           & 1.035                              & 0.000                             & 0.139    & 0.000 & 0.126    \\
+    {}                & AFNI         & -0.045                           & 1.282                              & 0.000                             & 0.373    & 0.000 & 0.355    \\
     \bottomrule
@@
 subject-level unthresholded t-statistic maps (mean over subjects).
-As for group-level maps, BT was consistently larger than WT (Wilcoxon
+As for group-level maps, the mean of differences in t-statistic maps was
+approximately zero centered in both BT and WT (t-test p~\textless~$10^{-4}$ for 13 subjects),
+and BT was consistently larger than WT (Wilcoxon
 signed-rank test and t-test p~\textless~$10^{-4}$ for 14 subjects).
@@
   \centering
-  \includegraphics[width=0.9\textwidth]{figures/plots/Bland-Altman/with-masking/unthresh-sbj04.png}
+  \includegraphics[width=0.9\textwidth]{figures/plots/Bland-Altman/with-masking/unthresh-sbj05.png}
   \caption{For subject with highest WT variability,
@@
 in both thresholded and unthresholded maps (see Table~\ref{table:pipeline-stats-fixed-mask}).
-The effect of masking was clearly appeared in BT by Bland-Altman plots showing that blob lines
+The effect of masking appeared in Bland-Altman plots by showing that blob lines
 were removed after fixing masks (Figure~\ref{fig:unthresh-maps-mask}-\textbf{A},\textbf{B}).
@@
     {}                & AFNI vs. SPM & 0.645                           & 1.428                              & 0.064                             & 1.489    & -0.065 & 1.206    \\
-    {Within Tool}     & FSL          & -0.025                           & 0.197                              & 0.012                             & 0.043    & 0.004 & 0.041    \\
-    {(WT)}            & SPM          & -0.002                           & 0.154                              & 0.000                             & 0.033    & 0.000 & 0.028    \\
-    {}                & AFNI         & -0.008                           & 0.208                              & 0.000                             & 0.078    & 0.000 & 0.072    \\
+    {Within Tool}     & FSL          & 0.012                           & 0.127                              & 0.000                             & 0.165    & 0.017 & 0.199    \\
+    {(WT)}            & SPM          & 0.002                           & 0.154                              & 0.000                             & 0.121    & 0.000 & 0.110    \\
+    {}                & AFNI         & -0.003                           & 0.3548                              & 0.003                             & 0.285    & 0.000 & 0.328    \\
     \bottomrule
@@
   \centering
-  \includegraphics[width=\textwidth]{figures/4-marked.pdf}
+  \includegraphics[width=\textwidth]{figures/5-marked.pdf}
   % \includegraphics[width=\textwidth]{figures/plots/wt-tstats/5-framed.png}

### 2022-02-21 computs.py

@@ -35,25 +35,25 @@ def compute_rel(f1, file_name=None, f2=None):
 
     # Compute relative diff in WT
     else:
-        sample_list = []
-        flag_ = False
-        for i in range(1, 11):
-            file_ = f1.replace('MCA', str(i))
-            file_ = nib.load(file_)
-            file_d = file_.get_fdata()
-            sample_list.append(file_d)
-            if flag_:
-                mask = np.where((file_d == 0) & (tmp == 0), np.nan, 1)
-            flag_ = True
-            tmp = file_d
-        diff_list = []
-        for pair_ in list(itertools.combinations(sample_list, 2)):
-            f1_data = pair_[0]
-            f2_data = pair_[1]
-            diff_list.append((f1_data - f2_data))
-        diff_avg = np.mean(diff_list, axis=0)
-        img_rel = np.where((mask == 1), diff_avg, np.nan)
-        nft_img_rel = nib.Nifti1Image(img_rel, file_.affine, header=file_.header)
+        file_ = f1.replace('MCA', '1')
+        file_ = nib.load(file_)
+        f1_data = file_.get_fdata()
+        max_ = np.zeros(f1_data.shape)
+        for pair_ in list(itertools.combinations(range(1, 11), 2)):
+            file_n1 = f1.replace('MCA', str(pair_[0]))
+            file_ = nib.load(file_n1)
+            f1_data = file_.get_fdata()
+            file_n2 = f1.replace('MCA', str(pair_[1]))
+            file_ = nib.load(file_n2)
+            f2_data = file_.get_fdata()
+            img_rel = (f1_data - f2_data)
+            img_rel = np.where((f1_data == 0) & (f2_data == 0), np.nan, img_rel)
+            if np.nanstd(img_rel) > np.nanstd(max_):
+                max_ = img_rel
+                f1_name = file_n1
+                f2_name = file_n2
+        print('{} and {} at {}'.format(f1_name, f2_name, file_name))
+        nft_img_rel = nib.Nifti1Image(max_, file_.affine, header=file_.header)
         nib.save(nft_img_rel, os.path.join('data/diff/', '{}-rel.nii.gz'.format(file_name)))
         return nft_img_rel
 
@@ -135,25 +135,7 @@ def var_between_fuzzy(mca_results):
[…]

### 2022-04-12 paper/paper.tex
Update small changes

@@
   approached between-tool software variability in some regions for some subjects.
-  Interestingly, within- and between-tool software variabilities appeared moderately
-  correlated ($r > 0.5$ across voxels) in all tested conditions. 
+  Interestingly, the brain masking instability was triggered by both within- and between-tool
+  software variabilities.
   % Finally, we found that
@@
 For both group- and subject-level analyses, between- and within-tool
-software variabilities were found to be moderately correlated. Even though
-both types of variability are different in nature, this result suggests
+software variabilities similarly trigger the brain masking instability.
+Even though both types of variability are different in nature, this result suggests
 that in some cases they may have a common cause that might be related to
-the conditioning of fMRI analysis in a specific dataset. Along these lines,
-in some regions, instabilities of similar magnitude may be triggered by
-small numerical perturbations, model variations, or implementation
-differences. For instance, high-motion datasets may trigger between- and
-within-tool instabilities similarly.
+the conditioning of fMRI analysis in a specific dataset. 
+% Along these lines,
+% in some regions, instabilities of similar magnitude may be triggered by
+% small numerical perturbations, model variations, or implementation
+% differences. For instance, high-motion datasets may trigger between- and
+% within-tool instabilities similarly.
 

### 2022-04-12 paper/paper.tex
Minor revisions

@@
   However, the relationship between within-tool and between-tool software
-  variabilities for a given analysis is unclear. Within-tool variability has multiple origins
-  including variations in algorithms, parametrizations, and
-  low-level software dependencies. We focus on the effects resulting from software dependencies --- 
-  in particular due to operating system and other infrastructural updates --- as they often remain unnoticed 
-  while algorithms and parametrizations are more controllable.
- We extended
-  a previous comparison of fMRI analysis software libraries (namely FSL,
-  AFNI, and SPM) and related the observed differences to within-tool
-  software variability simulated using Monte-Carlo arithmetic. In group
-  analyses, we found that
-  between-tool software variability was consistently larger than within-tool software
-  variability. In subject analyses, within-tool software variability
-  approached between-tool software variability in some regions for some subjects.
-  Interestingly, the brain masking instability was triggered by both within- and between-tool
-  software variabilities.
-  % Finally, we found that
-  % within-tool software variability had effects of similar magnitude to between-tool
-  % software variability when simulated from numerical perturbations introduced at the virtual
-  % precision of 17 bits --- a high but plausible level of uncertainty for
-  % scientific computing libraries. \tristan{Revise from latest results}
-  % %  \gk{What does this mean? They only have
-  % precision of 17, or they use data that is 17 bits?}
-  %  \yohan{agree with Greg, it is unclear and a strong assertion!}
-  %  \camille{It is not entirely clear to me how this relates to the previous
-  %  conclusions that BT variability was consistently larger than machine
-  %  error} \tristan{Better?}
-  Our findings motivate the continued investigation of within-tool
-  software variability in neuroimaging, and position it as a possible proxy
-  for studying some of the arbitrary between-tool software variations for
-  subject-level analyses.
+  variabilities for a given analysis is unclear. Within-tool variability has
+  multiple origins including variations in algorithms, parametrizations, and
+  low-level software dependencies. We focus on the effects resulting from
+  software dependencies --- in particular due to operating system and other
+  infrastructural updates --- as they often remain unnoticed while algorithms
+  and parametrizations are more controllable. We extended a previous
+  comparison of fMRI analysis software libraries (namely FSL, AFNI, and SPM)
+  and related the observed differences to within-tool software variability
+  simulated using Monte-Carlo arithmetic. In group analyses, we found that
+  between-tool software variability was consistently larger than within-tool
+  software variability. In subject analyses, between-tool software
+  variability also dominated within-tool variability overall, however,
+  within-tool software variability approached between-tool software
+  variability in some regions for some subjects. Interestingly, the brain
+  masking instability was triggered by both within- and between-tool software
+  variabilities. Our findings motivate the continued investigation of
+  within-tool software variability in neuroimaging.
   \end{abstract}
@@
 strictly identical results with the same MD5 checksums. For SPM, the files (as measured by MD5 checksums) were different but differences
-were visually unnoticeable. For AFNI, the overall patterns of activations were preseved
-, however, differences were noticeable visually.
+were visually unnoticeable. For AFNI, the overall patterns of activations were preserved, however, differences were noticeable visually.
 The observed differences remained small (see Supplemental Material~\ref{sec:supp-repro}), and might be due to the use of
@@
 However, for each tool pair (A, B), BT variability was
-significantly larger than WT in tool A or B,
-with an order of magnitude larger standard deviation of differences 
-in both thresholded and unthresholded maps, 
+significantly larger than WT in tool A or B
 (Wilcoxon signed-rank test and t-test p~\textless~$10^{-5}$).
@@
 we also computed results in the intersection of maps and found that the brain masking instability was
-triggered by both BT and WT.
-BT and WT variabilities were consistently smaller than ones with the fixed masks
-in both thresholded and unthresholded maps (see Table~\ref{table:pipeline-stats-fixed-mask}).
+triggered by both BT and WT (Table~\ref{table:pipeline-stats-fixed-mask}).
+BT and WT variabilities were generally smaller than the ones obtained previously (Table~\ref{table:pipeline-stats})
+in both thresholded and unthresholded maps.
 The effect of masking appeared in Bland-Altman plots by showing that blob lines
@@
 software variabilities similarly trigger the brain masking instability.
-Even though both types of variability are different in nature, this result suggests
-that in some cases they may have a common cause that might be related to
-the conditioning of fMRI analysis in a specific dataset. 
-% Along these lines,
-% in some regions, instabilities of similar magnitude may be triggered by
-% small numerical perturbations, model variations, or implementation
-% differences. For instance, high-motion datasets may trigger between- and
-% within-tool instabilities similarly.
-
-Therefore, numerical stability may be a suitable proxy to study
-between-tool variability. This speculation might be of practical value to
-address software variability at large given that within-tool variability
-measurements can be automated while between-tool ones require substantial
-human intervention. Indeed, within-tool variability can be simulated
-irrespective of the particular analysis done, and with important sample
-sizes, while between-tool variability requires re-implementing analysis in
-different toolboxes and is limited to a few samples. Automated multi-verse
-frameworks could be developed to assess and possibly reduce within-tool
-variability as a pre-requisite for between-tool variability.
-
-% \tristan{This paragraph needs to be revised as we can't conclude that from the data we have}
-% The finding that numerical variability approached tool variability at the
-% virtual precision of t=17~bits is interesting too. Indeed, while machine
-% error generally introduces perturbations on the least-significant bit of the mantissa,
-% ones 
-% % \yohan{I would not mix virtual precision and error measurement}\tristan{better?}
-% common scientific software dependencies introduce larger perturbations
-% .
-% For instance, SciPy's 2D spline interpolation was recently found to be
-% precise up to 10~bits~\cite{pytracer}
-% and its replacement by a more precise implementation might therefore introduce
-% numerical perturbations leading to errors in the range of between-tool
-% variability. Such high errors
-% might be triggered by updates in operating systems, Python, MATLAB, and
-% other software dependencies.
-% \yohan{The sentence is ok, just a picky point: the SciPy result you cited
-% describes the precision of the spline implementation while at the end
-% you're talking differences due to updates. The nuance is that two versions
-% may have a similar precision and so when you udpate your system, the
-% difference between both versions is low but each version can be
-% ill-conditionned.}
-% \tristan{better?}
+Even though both types of variability are different in nature, this result
+suggests that in some cases they may have a common cause that might be
+related to the conditioning of fMRI analysis in a specific dataset. In such
+cases, numerical stability may be a suitable proxy to study between-tool
+variability, which could be of practical value.
 
@@
 
-In conclusion, our results motivate further numerical stability
-investigations in fMRI analyses. Pipeline-level analyses could be conducted
-to identify specific components that contribute to numerical variability,
-and if possible correct them accordingly. Besides, in cases where
-instability is inherent to the analysis, sampling results distributions
-through numerical perturbations might improve stability, as explored
-in~\cite{kiar2021data}. Finally, pipeline-specific statistical corrections
-might be envisaged to account for within-tool software variability.
+In conclusion, between-tool variability clearly dominates within-tool
+variability overall. However, within-tool variability still reaches
+magnitudes that may affect scientific conclusions derived from data
+analyses --- in particular in subject analyses and in group analyses
+that rely on thresholded maps. Moreover, sources of variability such as brain
+masking affect both within-tool and between-tool
+variability: in these specific cases only, within-tool variability may be
+used as a proxy to between-tool variability. These findings motivate
+further investigations of within-tool varability in fMRI analyses.
 

### 2022-06-06 paper/paper.tex

@@
   masking instability was triggered by both within- and between-tool software
-  variabilities. Our findings motivate the continued investigation of
+  variabilities \camille{I did not quite understand what `masking instability` meant here, maybe ``differences observed in brain masks?''}. Our findings motivate the continued investigation of
   within-tool software variability in neuroimaging.
@@
 Mean of differences in t-statistic maps was centered around zero for BT and WT
-in both thresholded and unthresholded maps (t-test p~\textless~$10^{-5}$).
+in both thresholded and unthresholded maps (t-test p~\textless~$10^{-5}$ \camille{I don't understand the `centered around zero' and then the very small p-value, is this contradictory?
+} \camille{I am no statistician but I think we cannot perform t-test to test for differences in t-statistic.
+To perform a parametric test we would have to transform to z-stat or to go back to the contrast values.
+This reminds me of how we do tests in meta-analysis, more on page 3 of (unpublished) \url{https://github.com/cmaumet/zmeta_paper/blob/master/ibma_zmeta_journal.pdf}.
+I think the right thing to do here would be to use some version of Stouffer, if that's of interest I am happy  to chat about it?
+Hopefully this should not be too cumbersome, mainly T to Z transfo followed by some averaging and then testing against a normal distribution.
+Let me know!} ).
 However, for each tool pair (A, B), BT variability was
@@
   \includegraphics[width=.9\textwidth]{figures/rel-corr-unthresh-with-masking.png}
-  \caption{Voxel-wise comparison of group-level differences in BT and WT.}
+  \caption{Voxel-wise comparison of group-level differences in BT and WT. \camille{That's not obvious to me why are the bins bigger here compared to the B-A plots? }}
   \label{fig:unthresh-maps-corr}
@@
 and see Supplemental Material~\ref{sec:supp-worst-subject}).
+\camille{to me, this seems like an important result, is this big difference also visible on the thresholded maps of that subject?
+if so, that's just a thought of course, but I think that it could be very nice to also display the thresholded brain maps that correspond to the subject in question (similar to figure 4).
+This would be an important piece of info for brain imagers to understand the impact WT variability can have on their subject-level results.}
+
 % As for the group maps, BT and WT
@@
 active regions for both BT and WT. Except at the edges, BT remained consistently larger
-than WT (Figure~\ref{fig:thresh-maps}-\textbf{A},\textbf{B},\textbf{C} and Table~\ref{table:pipeline-stats}).
+than WT (Figure~\ref{fig:thresh-maps}-\textbf{A},\textbf{B},\textbf{C} and Table~\ref{table:pipeline-stats}) \camille{I personnally really like Fig.
+4!
+It shows very nicely how instable some clusters are and this is close to what brain imagers are used to.
+Maybe it could be in the main text instead of supplementary?}.
 
 Moreover, to compare the effects of BT and WT on the detection of activated
-regions, we measured WT instability and BT instability in each region of
+regions, we measured WT instability \camille{Could we describe in a sentence what WT instability mean in this context?
+i.e.
+when is it zero and when is it one.
+Sorry if I missed it!} and BT instability in each region of
 the HCP-MMP1.0 parcellation. The confusion matrices in
@@
 The effect of masking appeared in Bland-Altman plots by showing that blob lines
-were removed after fixing masks (Figure~\ref{fig:unthresh-maps-mask}-\textbf{A},\textbf{B}).
+were removed after fixing masks \camille{maybe change `fixing masks` into `when using the intersection mask for all maps` or something similar to avoid making it sound like what was done previously could have been fixed?.} (Figure~\ref{fig:unthresh-maps-mask}-\textbf{A},\textbf{B}).
 
